---
title: "Alternate normalization level analyses"
subtitle: "For review only"
author: "Kim Dill-McFarland"
date: '2018-10-11'
output:
  pdf_document: default
editor_options: 
  chunk_output_type: console
---
# Set seed
```{r}
set.seed(8325)
```

# Packages
```{r Load packages}
#Data manipulation
library(tidyverse)
#Beta-diversity analyses and plots
library(vegan)
```

# Data
## OTU tables
Rows are samples, columns are OTUs, data are given in counts
```{r message=FALSE}
#Fungi
OTUF = read_tsv("data/OTU_table/OTU.Fun.norm1000.txt") %>% rename(Sample=Samle) %>% separate(Sample, c("Sample", "amplicon"), sep=".Fun")
```

## Metadata by sample

* Animal: animal ID number
* Diet: calf starter (A), corn silage (B), mixture (C)
* AgeGroup: 2 weeks (2w), 4 weeks (4w), 8 weeks (8w), 1 year (1yr), 2 years (2yr)
* AgeExact: exact age in days on date of sampling
* Daysln: natural log of exact age
* TypeAll: feces (F), rumen liquids (L, no number for calves, 1-3 for day on repeated sampling for cows), rumen solids (S, no number for calves, 1-3 for day of consecutive sampling for cows)
* Type: feces (F), rumen liquids (L), rumen solids (S) (simplified to not include numbers for day)
* Sex: males (M), female (F)
* ScourSample: Y/N if sample was taken while calf was experiencing scours
* chao: Chao1 richness estimate
* shannon: Shannon's diversity estimate
* ADGKG: average daily gain in kg from 2 days of age to date of sampling
* Gain: total weight gain in kg relative to 2 day measuremnts
* TotalIntake: total supplement intake (g) from birth up to date of sampling
* ill.scour: total number of days ill with scours up to date of sampling
* ill.resp: total number of days ill with scours up to date of sampling
* ill.ketosis: total number of days ill with ketosis up to date of sampling
* ill.mastitis: total number of days ill with mastitis up to date of sampling
* ill.all: total number of days ill (any recorded illness) up to date of sampling
* MPE: milk production efficiency (energy corrected milk [ECM] / dry matter intake [DMI]) (kg/kg DMI)
* Effic: total energy output efficiency (MJ/kg DMI)
    
```{r message=FALSE}
#Fungi
metaF = read_tsv("data/metadata/Met.Fun2.txt", col_types = list(Animal = col_character()))
```

# Add Diet + Age variable
```{r}
#Fungi
metaF$DietAge = as.factor(paste(metaF$Diet, metaF$AgeGroup, sep=""))
```

# Average consecutive 1 & 2yr samples
To remove animal effects in some analyses, average the data from 1 and 2 year samples taken across 3 consecutive days.

## OTU tables
**Fungi**
```{r}
OTUFave = metaF %>%
  # Merge OTU with metadata needed for grouping 3 day replicates in OTU data
  select(Sample, Animal, AgeGroup, Type) %>% 
  right_join(OTUF, by="Sample") %>%
  select(-Sample) %>% 
  # average 3 day replicates for 1 and 2 yr rumen samples
  group_by(Animal, AgeGroup, Type) %>% 
  summarize_if(is.numeric, funs(mean)) %>%
  # round to whole numbers
  mutate_if(is.numeric, funs(round(., 0))) %>% 
  ungroup() %>% 
  # recreate Sample name as row name
  mutate(Sample = paste(Animal, AgeGroup, sep=".")) %>% 
  mutate(Sample = paste(Sample, Type, sep=".")) %>% 
  select(-Animal, -AgeGroup, -Type)
```

## Metadata by sample
**Fungi**
```{r}
metaFave = metaF %>%
  # average 3 day replicates for 1 and 2 yr rumen samples
  group_by(Animal, AgeGroup, Type, Diet) %>% 
  summarize_if(is.numeric, funs(mean)) %>%
  ungroup() %>% 
  # recreate Sample name as row name
  mutate(Sample = paste(Animal, AgeGroup, sep=".")) %>% 
  mutate(Sample = paste(Sample, Type, sep="."))
```

# BETA-DIVERSITY
# Average beta by age
## Fungi
**Bray-Curtis**
```{r}
BC.temp = OTUF %>%
  # Remove non-numeric labels
  select(-Sample,-amplicon) %>% 
  vegdist(method="bray") %>% 
  # Convert to data frame
  as.matrix() %>% 
  as.data.frame() %>% 
  # Label columns and rows by sample name
  mutate(Sample=OTUF$Sample) %>%
  rename_if(is.numeric, ~ OTUF$Sample) %>% 
  # Add metadata
  full_join(metaF[,c("Sample","Animal","AgeGroup","Type")], by="Sample") %>% 
  # Reformat to long
  gather(-Sample, -Animal, -AgeGroup, -Type, key="Sample2", value="BC") %>% 
  # Remove self comparisons
  dplyr::filter(BC > 0) %>% 
  # Calculate mean and stdev
  group_by(AgeGroup, Type) %>% 
  summarize(mean=mean(BC, na.rm=TRUE), sd=sd(BC, na.rm=TRUE)) %>% 
  mutate(amplicon="Fun", metric="BC")
```

**Jaccard**
```{r}
J.temp = OTUF %>%
  # Remove non-numeric labels
  select(-Sample, -amplicon) %>% 
  vegdist(method="jaccard") %>% 
  # Convert to data frame
  as.matrix() %>% 
  as.data.frame() %>% 
  # Label columns and rows by sample name
  mutate(Sample=OTUF$Sample) %>%
  rename_if(is.numeric, ~ OTUF$Sample) %>% 
  # Add metadata
  full_join(metaF[,c("Sample","Animal","AgeGroup","Type")], by="Sample") %>% 
  # Reformat to long
  gather(-Sample, -Animal, -AgeGroup, -Type, key="Sample2", value="J") %>% 
  # Remove self comparisons
  dplyr::filter(J > 0) %>% 
  # Calculate mean and stdev
  group_by(AgeGroup, Type) %>% 
  summarize(mean=mean(J, na.rm=TRUE), sd=sd(J, na.rm=TRUE)) %>% 
  mutate(amplicon="Fun", metric="J")
```

**Combine**
```{r eval=FALSE}
write.table(rbind(BC.temp, J.temp), "results/beta_div/F.beta.txt", row.names=FALSE, sep="\t")
```

## Combine all
```{r message=FALSE}
beta.ave = read_tsv("results/beta_div/A.beta.txt") %>%
  rbind(read_tsv("results/beta_div/B.beta.txt")) %>% 
  rbind(read_tsv("results/beta_div/F.beta.txt")) %>%
  gather(mean:sd, key=stat, value=value) %>% 
  mutate(group=paste(AgeGroup, stat)) %>% 
  select(-stat,-AgeGroup) %>% 
  group_by(Type, amplicon, metric) %>% 
  spread(key=group, value=value)
```

# Model 1: Diet and age
Diet * AgeGroup

* Permutational ANOVA (PERMANOVA) followed by pairwise PERMANOVA with FDR correction (Benjamini-Hochberg method) if significant

## Fungi
### Fecal
Subset data to fecal samples
```{r}
OTU.temp = metaF %>%
  select(Sample, Type, Diet, AgeGroup, DietAge) %>% 
  filter(Type == "F") %>% 
  left_join(OTUF, by="Sample") %>% 
  arrange(Sample)

meta.temp = metaF %>%
  filter(Type == "F")%>% 
  arrange(Sample)
```

#### PERMANOVA
Bray-Curtis (structure)
```{r}
adonis(OTU.temp[, grepl("Otu", colnames(OTU.temp))] ~ Diet*AgeGroup, data=meta.temp, method="bray", permutations=1000)
```

Jaccard (composition)
```{r}
adonis(OTU.temp[, grepl("Otu", colnames(OTU.temp))] ~ Diet*AgeGroup, data=meta.temp, method="jaccard", permutations=1000)
```

### Rumen liquids
Subset data to liquid samples
```{r}
OTU.temp = metaF %>%
  select(Sample, Type, Diet, AgeGroup, DietAge) %>% 
  filter(Type == "L") %>% 
  left_join(OTUF, by="Sample") %>% 
  arrange(Sample)

meta.temp = metaF %>%
  filter(Type == "L") %>% 
  arrange(Sample)
```

#### PERMANOVA
Bray-Curtis (structure)
```{r}
adonis(OTU.temp[, grepl("Otu", colnames(OTU.temp))] ~ Diet*AgeGroup, data=meta.temp, method="bray", permutations=1000)
```

Jaccard (composition)
```{r}
adonis(OTU.temp[, grepl("Otu", colnames(OTU.temp))] ~ Diet*AgeGroup, data=meta.temp, method="jaccard", permutations=1000)
```

### Rumen solids
Subset data to solid samples
```{r}
OTU.temp = metaF %>%
  select(Sample, Type, Diet, AgeGroup, DietAge) %>% 
  filter(Type == "S") %>% 
  left_join(OTUF, by="Sample") %>% 
  arrange(Sample)

meta.temp = metaF %>%
  filter(Type == "S") %>% 
  arrange(Sample)
```

#### PERMANOVA
Bray-Curtis (structure)
```{r}
adonis(OTU.temp[, grepl("Otu", colnames(OTU.temp))] ~ Diet*AgeGroup, data=meta.temp, method="bray", permutations=1000)
```

Jaccard (composition)
```{r}
adonis(OTU.temp[, grepl("Otu", colnames(OTU.temp))] ~ Diet*AgeGroup, data=meta.temp, method="jaccard", permutations=1000)
```

# Model 2: Milk production efficiency
2yr samples only, averaged 3 consecutive rumen samples to avoid animal effects  
Effic + MPE

* Permutational ANOVA (PERMANOVA) followed by pairwise PERMANOVA with FDR correction (Benjamini-Hochberg method) if significant

## Fungi
### Fecal
Subset data to 2yr fecal samples
```{r}
## Metadata
meta.temp = metaFave %>%
  filter(Type == "F" & AgeGroup == "2yr")
```

```{r}
OTU.temp = metaFave %>%
  # Combine with metadata
  select(Sample, AgeGroup, Type) %>% 
  right_join(OTUFave, by="Sample") %>%
  # Filter to only 2 yr fecal samples
  filter(Type == "F" & AgeGroup == "2yr") %>% 
  # Remove metadata variables
  select(-Sample, -AgeGroup, -Type)
```

**PERMANOVA**  
Bray-Curtis (structure)
```{r}
adonis(OTU.temp ~ Effic+MPE, data=meta.temp, method="bray", permutations=1000)
```

Jaccard (composition)
```{r}
adonis(OTU.temp ~ Effic+MPE, data=meta.temp, method="jaccard", permutations=1000)
```

### Rumen liquids
Subset data to 2yr rumen liquid samples
```{r}
## Metadata
meta.temp = metaFave %>%
  # Efficieny 
  filter(Type == "L" & AgeGroup == "2yr")
```

```{r}
OTU.temp = metaFave %>%
  # Combine with metadata
  select(Sample, AgeGroup, Type) %>% 
  right_join(OTUFave, by="Sample") %>%
  # Filter to only 2 yr fecal samples
  filter(Type == "L" & AgeGroup == "2yr") %>% 
  # Remove metadata variables
  select(-Sample, -AgeGroup, -Type)
```

**PERMANOVA**  
Bray-Curtis (structure)
```{r}
adonis(OTU.temp ~ Effic+MPE, data=meta.temp, method="bray", permutations=1000)
```

Jaccard (composition)
```{r}
adonis(OTU.temp ~ Effic+MPE, data=meta.temp, method="jaccard", permutations=1000)
```

### Rumen solids
Subset data to 2yr rumen solid samples
```{r}
## Metadata
meta.temp = metaFave %>%
  # Efficieny 
  filter(Type == "S" & AgeGroup == "2yr")
```

```{r}
OTU.temp = metaFave %>%
  # Combine with metadata
  select(Sample, AgeGroup, Type) %>% 
  right_join(OTUFave, by="Sample") %>%
  # Filter to only 2 yr fecal samples
  filter(Type == "S" & AgeGroup == "2yr") %>% 
  # Remove metadata variables
  select(-Sample, -AgeGroup, -Type)
```

**PERMANOVA**  
Bray-Curtis (structure)
```{r}
adonis(OTU.temp ~ Effic+MPE, data=meta.temp, method="bray", permutations=1000)
```

Jaccard (composition)
```{r}
adonis(OTU.temp ~ Effic+MPE, data=meta.temp, method="jaccard", permutations=1000)
```

# ALPHA-DIVERSITY
# Model 1: Diet and age
Diet * AgeGroup

* ANOVA

## Fungi
### Fecal
#### Shannon's diversity
**ANOVA**
```{r}
metaF %>%
  filter(Type == "F") %>% 
  aov(shannon ~ Diet*AgeGroup, .) %>% 
  summary()
```

### Rumen liquids
#### Shannon's diversity
**ANOVA**
```{r}
metaF %>%
  filter(Type == "L") %>% 
  aov(shannon ~ Diet*AgeGroup, .) %>% 
  summary()
```

### Rumen solids
#### Shannon's diversity
**ANOVA**
```{r}
metaF %>%
  filter(Type == "S") %>% 
  aov(shannon ~ Diet*AgeGroup, .) %>% 
  summary()
```

# Model 2: Milk production efficiency
2yr samples only, averaged 3 consecutive rumen samples to avoid animal effects  
Effic + MPE

* ANOVA followed by pairwise TukeyHSD if significant

## Fungi
### Fecal
#### Shannon's diversity
**ANOVA**
```{r}
metaFave %>%
  filter(Type == "F" & AgeGroup == "2yr") %>% 
  aov(shannon ~ Effic+MPE, .) %>% 
  summary()
```

### Rumen liquids
#### Shannon's diversity
**ANOVA**
```{r}
metaFave %>%
  filter(Type == "L" & AgeGroup == "2yr") %>% 
  aov(shannon ~ Effic+MPE, .) %>% 
  summary()
```

### Rumen solids
#### Shannon's diversity
**ANOVA**
```{r}
metaFave %>%
  filter(Type == "S" & AgeGroup == "2yr") %>% 
  aov(shannon ~ Effic+MPE, .) %>% 
  summary()
```