---
title: "Data manipulation"
subtitle: "GS01 Calf diets"
author: "Kim Dill-McFarland"
date: "version `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_float:
      collapsed: no
editor_options: 
  chunk_output_type: console
---
```{r}
sessionInfo()
```

# Set seed
```{r}
set.seed(8325)
```

# Packages
```{r Load packages}
#Data manipulation
library(tidyverse)
```

# Data
## OTU tables
Rows are samples, columns are OTUs, data are given in counts
```{r message=FALSE}
#Archaea
OTUA = read_tsv("data/OTU_table/OTU.Arc.norm100.txt")

#Bacteria
OTUB = read_tsv("data/OTU_table/OTU.Bact.norm9000.txt")

#Fungi
OTUF = read_tsv("data/OTU_table/OTU.Fun.norm250.txt")
```

## Metadata by sample

* Animal: animal ID number
* Diet: calf starter (A), corn silage (B), mixture (C)
* AgeGroup: 2 weeks (2w), 4 weeks (4w), 8 weeks (8w), 1 year (1yr), 2 years (2yr)
* AgeExact: exact age in days on date of sampling
* Daysln: natural log of exact age
* TypeAll: feces (F), rumen liquids (L, no number for calves, 1-3 for day on repeated sampling for cows), rumen solids (S, no number for calves, 1-3 for day of consecutive sampling for cows)
* Type: feces (F), rumen liquids (L), rumen solids (S) (simplified to not include numbers for day)
* Sex: males (M), female (F)
* ScourSample: Y/N if sample was taken while calf was experiencing scours
* chao: Chao1 richness estimate
* shannon: Shannon's diversity estimate
* ADGKG: average daily gain in kg from 2 days of age to date of sampling
* Gain: total weight gain in kg relative to 2 day measuremnts
* TotalIntake: total supplement intake (g) from birth up to date of sampling
* ill.scour: total number of days ill with scours up to date of sampling
* ill.resp: total number of days ill with scours up to date of sampling
* ill.ketosis: total number of days ill with ketosis up to date of sampling
* ill.mastitis: total number of days ill with mastitis up to date of sampling
* ill.all: total number of days ill (any recorded illness) up to date of sampling
* MPE: milk production efficiency (energy corrected milk [ECM] / dry matter intake [DMI]) (kg/kg DMI)
* Effic: total energy output efficiency (MJ/kg DMI)
    
```{r message=FALSE}
#Archaea
metaA = read_tsv("data/metadata/Met.Arc.txt", col_types = list(Animal = col_character()))

#Bacteria
metaB = read_tsv("data/metadata/Met.Bact.txt", col_types = list(Animal = col_character()))

#Fungi
metaF = read_tsv("data/metadata/Met.Fun.txt", col_types = list(Animal = col_character()))
```

## Supplement intake

* Animal: animal ID number
* Diet: calf starter (A), corn silage (B), mixture (C)
* Day: age in days
* Intake: supplement intake in g (feed offered - feed refused)

```{r message=FALSE}
intake = read_tsv("data/metadata/feed_intake.txt", col_types = list(Animal = col_character()))
```

## SCOA in mM

Rows are samples, columns are short chain organic acids (SCOAs), data are given as concentrations (mM). ETOH = ethanol, iVal.2MB = isovalerate + 2-methylbutyrate

```{r message=FALSE}
SCOA = read_tsv("data/SCOA/SCOA.txt")
```

## VFA in mF

Rows are samples, columns are volatile fatty acids (VFas), data are given as concentrations (mM) or molar fractions (mF), **except**

* Diet: calf starter (A), corn silage (B), mixture (C)
* AgeGroup: 8 weeks (8w), 1 year (1yr), 2 years (2yr)
* TypeAll: (L, no number for calves, 1-3 for day on repeated sampling for cows)

isoval_2MB = isovalerate + 2-methylbutyrate, branch = isobutyrate + isovalerate + 2-methylbutyrate

```{r message=FALSE}
VFA = read_tsv("data/SCOA/VFA_mF.txt")
```

## Weight

* Animal: animal ID number
* Diet: calf starter (A), corn silage (B), mixture (C)
* AgeExact: exact age in days on date of weighing
* Wt: weight in kg
* Age_norm: AgeExact - 2 in order to normalize to number of days ellapsed since first weighing at 2 days of age
* Wt_norm: weight gained since day 2

```{r message=FALSE}
weight = read_tsv("data/metadata/weight_gain.txt", col_types = list(Animal = col_character()))
```

# Add Diet + Age variable
```{r}
#Archaea
metaA$DietAge = as.factor(paste(metaA$Diet, metaA$AgeGroup, sep=""))

#Bacteria
metaB$DietAge = as.factor(paste(metaB$Diet, metaB$AgeGroup, sep=""))

#Fungi
metaF$DietAge = as.factor(paste(metaF$Diet, metaF$AgeGroup, sep=""))
```

# Average consecutive 1 & 2yr samples
To remove animal effects in some analyses, average the data from 1 and 2 year samples taken across 3 consecutive days.

## OTU tables
**Archaea**
```{r}
OTUAave = metaA %>%
  # Merge OTU with metadata needed for grouping 3 day replicates in OTU data
  select(Sample, Animal, AgeGroup, Type) %>% 
  right_join(OTUA, by="Sample") %>%
  # average 3 day replicates for 1 and 2 yr rumen samples
  group_by(Animal, AgeGroup, Type) %>% 
  summarize_if(is.numeric, funs(mean)) %>%
  # round to whole numbers
  mutate_if(is.numeric, funs(round(., 0))) %>% 
  ungroup() %>% 
  # recreate Sample name as row name
  mutate(Sample = paste(Animal, AgeGroup, sep=".")) %>% 
  mutate(Sample = paste(Sample, Type, sep=".")) %>% 
  select(-Animal, -AgeGroup, -Type)
```

**Bacteria**
```{r eval=FALSE}
OTUBave = metaB %>%
  # Merge OTU with metadata needed for grouping 3 day replicates in OTU data
  select(Sample, Animal, AgeGroup, Type) %>% 
  right_join(OTUB, by="Sample") %>%
  # average 3 day replicates for 1 and 2 yr rumen samples
  group_by(Animal, AgeGroup, Type) %>% 
  summarize_if(is.numeric, funs(mean)) %>%
  # round to whole numbers
  mutate_if(is.numeric, funs(round(., 0))) %>% 
  ungroup() %>% 
  # recreate Sample name as row name
  mutate(Sample = paste(Animal, AgeGroup, sep=".")) %>% 
  mutate(Sample = paste(Sample, Type, sep=".")) %>% 
  select(-Animal, -AgeGroup, -Type)
```

```{r echo=FALSE, eval=FALSE}
# Write resulting table for bacteria
write.table(OTUBave, file="data/OTU_table/OTUBave.txt", sep="\t")
```

```{r message=FALSE, echo=FALSE}
# Read in resulting table for bacteria as-rerunning clean-up is lengthy
OTUBave = read.table("data/OTU_table/OTUBave.txt", sep="\t", header=TRUE, row.names=1)
```

**Fungi**
```{r}
OTUFave = metaF %>%
  # Merge OTU with metadata needed for grouping 3 day replicates in OTU data
  select(Sample, Animal, AgeGroup, Type) %>% 
  right_join(OTUF, by="Sample") %>%
  select(-Sample) %>% 
  # average 3 day replicates for 1 and 2 yr rumen samples
  group_by(Animal, AgeGroup, Type) %>% 
  summarize_if(is.numeric, funs(mean)) %>%
  # round to whole numbers
  mutate_if(is.numeric, funs(round(., 0))) %>% 
  ungroup() %>% 
  # recreate Sample name as row name
  mutate(Sample = paste(Animal, AgeGroup, sep=".")) %>% 
  mutate(Sample = paste(Sample, Type, sep=".")) %>% 
  select(-Animal, -AgeGroup, -Type)
```

## Metadata by sample
**Archaea**
```{r}
metaAave = metaA %>%
  # average 3 day replicates for 1 and 2 yr rumen samples
  group_by(Animal, AgeGroup, Type, Diet) %>% 
  summarize_if(is.numeric, funs(mean)) %>%
  ungroup() %>% 
  # recreate Sample name as row name
  mutate(Sample = paste(Animal, AgeGroup, sep=".")) %>% 
  mutate(Sample = paste(Sample, Type, sep="."))
```

**Bacteria**
```{r}
metaBave = metaB %>%
  # average 3 day replicates for 1 and 2 yr rumen samples
  group_by(Animal, AgeGroup, Type, Diet) %>% 
  summarize_if(is.numeric, funs(mean)) %>%
  ungroup() %>% 
  # recreate Sample name as row name
  mutate(Sample = paste(Animal, AgeGroup, sep=".")) %>% 
  mutate(Sample = paste(Sample, Type, sep="."))
```

**Fungi**
```{r}
metaFave = metaF %>%
  # average 3 day replicates for 1 and 2 yr rumen samples
  group_by(Animal, AgeGroup, Type, Diet) %>% 
  summarize_if(is.numeric, funs(mean)) %>%
  ungroup() %>% 
  # recreate Sample name as row name
  mutate(Sample = paste(Animal, AgeGroup, sep=".")) %>% 
  mutate(Sample = paste(Sample, Type, sep="."))
```

## SCOA data
```{r}
SCOAave = SCOA %>%
  #Add Type column (L only)
  mutate(Type="L") %>% 
  dplyr::select(-Sample, -TypeAll) %>% 
  # average 3 day replicates for 1 and 2 yr rumen samples
  group_by(Animal, AgeGroup, Type, Diet) %>% 
  summarize_all(funs(mean)) %>%
  ungroup() %>% 
  # recreate Sample name as row name
  mutate(Sample = paste(Animal, AgeGroup, sep=".")) %>% 
  mutate(Sample = paste(Sample, Type, sep=".")) %>% 
  select(-Animal, -AgeGroup, -Type)
```

## VFA data
```{r}
VFAave = VFA %>%
  #Add Type column (L only)
  mutate(Type="L") %>% 
  dplyr::select(-Sample, -TypeAll) %>% 
  # average 3 day replicates for 1 and 2 yr rumen samples
  group_by(Animal, AgeGroup, Type, Diet) %>% 
  summarize_all(funs(mean)) %>%
  ungroup() %>% 
  # recreate Sample name as row name
  mutate(Sample = paste(Animal, AgeGroup, sep=".")) %>% 
  mutate(Sample = paste(Sample, Type, sep=".")) %>% 
  select(-Animal, -AgeGroup, -Type)
```

# Save environment
```{r}
save.image("GS01.data.RData")
```