---
title: "Animal metric analyses"
subtitle: "GS01 Calf diets"
author: "Kim Dill-McFarland"
date: "version `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_float:
      collapsed: no
editor_options: 
  chunk_output_type: console
---
<script>
$(document).ready(function() {
  $items = $('div#TOC li');
  $items.each(function(idx) {
    num_ul = $(this).parentsUntil('#TOC').length;
    $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
  });
});
</script>

# Load data
Created in `1_data_manip.Rmd`
```{r}
set.seed(8325)
load("GS01.data.RData")
```

# Packages
```{r Load packages}
#Data manipulation
library(tidyverse)
#Repeated measure linear models
library(lme4)
library(lmerTest)
```

# Supplement intake
Repeated measures linear model of log supplement intake over time.
```{r}
#Remove 0s since will break log transformation
intake[intake == 0] <- NA

#Repeated measures
anova(lmer(log(Intake) ~ Diet + (1+Day|Animal), data=intake))
```

# Weight gain
## Model 1: Diet and age
Age_norm + Age_norm:Diet + (0+Age_norm|Animal)

* Repeated measure linear model weight gain over time
* Relative to day 2
    - Wt_norm = weight - weight at 2 days
    - Age_norm = age in days - 2
* Separate models for calfs (0-8 wk) and cows (8 wk - 2 yrs) as growth occurs more rapidly in young animals.

### Calves
```{r warning=FALSE}
summary(lmer(Wt_norm ~ Age_norm + Age_norm:Diet + (0+Age_norm|Animal), data = weight[weight$Age_norm<100,]))

anova(lmer(Wt_norm ~ Age_norm + Age_norm:Diet + (0+Age_norm|Animal), data = weight[weight$Age_norm<100,]))
```

### Cows
```{r warning=FALSE}
summary(lmer(Wt_norm ~ Age_norm + Age_norm:Diet + (0+Age_norm|Animal), data = weight[weight$Age_norm>100,]))

anova(lmer(Wt_norm ~ Age_norm + Age_norm:Diet + (0+Age_norm|Animal), data = weight[weight$Age_norm>100,]))
```

# Production metrics
## Model 4: Diet only
Diet

* ANOVA

### Milk production efficiency (MPE)
Energy-corrected milk (ECM) production per kg dry matter intake (DMI).

```{r}
metaBave %>% 
  filter(Type == "F" & AgeGroup == "2yr") %>% 
  aov(MPE ~ Diet, .) %>% 
  summary()
```

### Overall energy efficiency
Efficiency was defined as total energy output (MJ) of the five energetic demands (*i.e.* ECM, weight maintenance, weight change, gestation, mastitis) per kg of DMI.

```{r}
metaBave %>% 
  filter(Type == "F" & AgeGroup == "2yr") %>% 
  aov(Effic ~ Diet, .) %>% 
  summary()
```
