---
title: "Alpha-diversity analyses"
subtitle: "GS01 Calf diets"
author: "Kim Dill-McFarland"
date: "version `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_float:
      collapsed: no
editor_options: 
  chunk_output_type: console
---
<script>
$(document).ready(function() {
  $items = $('div#TOC li');
  $items.each(function(idx) {
    num_ul = $(this).parentsUntil('#TOC').length;
    $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
  });
});
</script>

# Load data
Created in `1_data_manip.Rmd`
```{r}
set.seed(8325)
load("GS01.data.RData")
```

# Packages
```{r Load packages}
#Data manipulation
library(tidyverse)
```

# Model 1: Diet and age
Diet * AgeGroup

* ANOVA followed by pairwise TukeyHSD if significant

## Archaea
### Fecal
#### Shannon's diversity
**ANOVA**
```{r}
metaA %>%
  filter(Type == "F") %>% 
  aov(shannon ~ Diet*AgeGroup, .) %>% 
  summary()
```

**TukeyHSD**
```{r}
metaA %>%
  filter(Type == "F") %>% 
  aov(shannon ~ Diet*AgeGroup, .) %>% 
  TukeyHSD()
```

#### Chao richness
**ANOVA**
```{r}
metaA %>%
  filter(Type == "F") %>% 
  aov(chao ~ Diet*AgeGroup, .) %>% 
  summary()
```

**TukeyHSD**
```{r}
metaA %>%
  filter(Type == "F") %>% 
  aov(chao ~ Diet*AgeGroup, .) %>% 
  TukeyHSD()
```

### Rumen liquids
#### Shannon's diversity
**ANOVA**
```{r}
metaA %>%
  filter(Type == "L") %>% 
  aov(shannon ~ Diet*AgeGroup, .) %>% 
  summary()
```

**TukeyHSD**
```{r}
metaA %>%
  filter(Type == "L") %>% 
  aov(shannon ~ Diet*AgeGroup, .) %>% 
  TukeyHSD()
```

#### Chao richness
**ANOVA**
```{r}
metaA %>%
  filter(Type == "L") %>% 
  aov(chao ~ Diet*AgeGroup, .) %>% 
  summary()
```

**TukeyHSD**
```{r}
metaA %>%
  filter(Type == "L") %>% 
  aov(chao ~ Diet*AgeGroup, .) %>% 
  TukeyHSD()
```

### Rumen solids
#### Shannon's diversity
**ANOVA**
```{r}
metaA %>%
  filter(Type == "S") %>% 
  aov(shannon ~ Diet*AgeGroup, .) %>% 
  summary()
```

**TukeyHSD**
```{r}
metaA %>%
  filter(Type == "S") %>% 
  aov(shannon ~ Diet*AgeGroup, .) %>% 
  TukeyHSD()
```

#### Chao richness
**ANOVA**
```{r}
metaA %>%
  filter(Type == "S") %>% 
  aov(chao ~ Diet*AgeGroup, .) %>% 
  summary()
```

**TukeyHSD**
```{r}
metaA %>%
  filter(Type == "S") %>% 
  aov(chao ~ Diet*AgeGroup, .) %>% 
  TukeyHSD()
```

## Bacteria
### Fecal
#### Shannon's diversity
**ANOVA**
```{r}
metaB %>%
  filter(Type == "F") %>% 
  aov(shannon ~ Diet*AgeGroup, .) %>% 
  summary()
```

**TukeyHSD**
```{r}
metaB %>%
  filter(Type == "F") %>% 
  aov(shannon ~ Diet*AgeGroup, .) %>% 
  TukeyHSD()
```

#### Chao richness
**ANOVA**
```{r}
metaB %>%
  filter(Type == "F") %>% 
  aov(chao ~ Diet*AgeGroup, .) %>% 
  summary()
```

**TukeyHSD**
```{r}
metaB %>%
  filter(Type == "F") %>% 
  aov(chao ~ Diet*AgeGroup, .) %>% 
  TukeyHSD()
```

### Rumen liquids
#### Shannon's diversity
**ANOVA**
```{r}
metaB %>%
  filter(Type == "L") %>% 
  aov(shannon ~ Diet*AgeGroup, .) %>% 
  summary()
```

**TukeyHSD**
```{r}
metaB %>%
  filter(Type == "L") %>% 
  aov(shannon ~ Diet*AgeGroup, .) %>% 
  TukeyHSD()
```

#### Chao richness
**ANOVA**
```{r}
metaB %>%
  filter(Type == "L") %>% 
  aov(chao ~ Diet*AgeGroup, .) %>% 
  summary()
```

**TukeyHSD**
```{r}
metaB %>%
  filter(Type == "L") %>% 
  aov(chao ~ Diet*AgeGroup, .) %>% 
  TukeyHSD()
```

### Rumen solids
#### Shannon's diversity
**ANOVA**
```{r}
metaB %>%
  filter(Type == "S") %>% 
  aov(shannon ~ Diet*AgeGroup, .) %>% 
  summary()
```

**TukeyHSD**
```{r}
metaB %>%
  filter(Type == "S") %>% 
  aov(shannon ~ Diet*AgeGroup, .) %>% 
  TukeyHSD()
```

#### Chao richness
**ANOVA**
```{r}
metaB %>%
  filter(Type == "S") %>% 
  aov(chao ~ Diet*AgeGroup, .) %>% 
  summary()
```

**TukeyHSD**
```{r}
metaB %>%
  filter(Type == "S") %>% 
  aov(chao ~ Diet*AgeGroup, .) %>% 
  TukeyHSD()
```

## Fungi
### Fecal
#### Shannon's diversity
**ANOVA**
```{r}
metaF %>%
  filter(Type == "F") %>% 
  aov(shannon ~ Diet*AgeGroup, .) %>% 
  summary()
```

**TukeyHSD**
```{r}
metaF %>%
  filter(Type == "F") %>% 
  aov(shannon ~ Diet*AgeGroup, .) %>% 
  TukeyHSD()
```

#### Chao richness
**ANOVA**
```{r}
metaF %>%
  filter(Type == "F") %>% 
  aov(chao ~ Diet*AgeGroup, .) %>% 
  summary()
```

**TukeyHSD**
```{r}
metaF %>%
  filter(Type == "F") %>% 
  aov(chao ~ Diet*AgeGroup, .) %>% 
  TukeyHSD()
```

### Rumen liquids
#### Shannon's diversity
**ANOVA**
```{r}
metaF %>%
  filter(Type == "L") %>% 
  aov(shannon ~ Diet*AgeGroup, .) %>% 
  summary()
```

**TukeyHSD**
```{r}
metaF %>%
  filter(Type == "L") %>% 
  aov(shannon ~ Diet*AgeGroup, .) %>% 
  TukeyHSD()
```

#### Chao richness
**ANOVA**
```{r}
metaF %>%
  filter(Type == "L") %>% 
  aov(chao ~ Diet*AgeGroup, .) %>% 
  summary()
```

**TukeyHSD**
```{r}
metaF %>%
  filter(Type == "L") %>% 
  aov(chao ~ Diet*AgeGroup, .) %>% 
  TukeyHSD()
```

### Rumen solids
#### Shannon's diversity
**ANOVA**
```{r}
metaF %>%
  filter(Type == "S") %>% 
  aov(shannon ~ Diet*AgeGroup, .) %>% 
  summary()
```

**TukeyHSD**
```{r}
metaF %>%
  filter(Type == "S") %>% 
  aov(shannon ~ Diet*AgeGroup, .) %>% 
  TukeyHSD()
```

#### Chao richness
**ANOVA**
```{r}
metaF %>%
  filter(Type == "S") %>% 
  aov(chao ~ Diet*AgeGroup, .) %>% 
  summary()
```

**TukeyHSD**
```{r}
metaF %>%
  filter(Type == "S") %>% 
  aov(chao ~ Diet*AgeGroup, .) %>% 
  TukeyHSD()
```

# Model 2: Milk production efficiency
2yr samples only, averaged 3 consecutive rumen samples to avoid animal effects  
Effic + MPE

* ANOVA followed by pairwise TukeyHSD if significant

## Archaea
### Fecal
#### Shannon's diversity
**ANOVA**
```{r}
metaAave %>%
  filter(Type == "F" & AgeGroup == "2yr") %>% 
  aov(shannon ~ Effic+MPE, .) %>% 
  summary()
```

#### Chao richness
**ANOVA**
```{r}
metaAave %>%
  filter(Type == "F" & AgeGroup == "2yr") %>% 
  aov(chao ~ Effic+MPE, .) %>% 
  summary()
```

### Rumen liquids
#### Shannon's diversity
**ANOVA**
```{r}
metaAave %>%
  filter(Type == "L" & AgeGroup == "2yr") %>% 
  aov(shannon ~ Effic+MPE, .) %>% 
  summary()
```

#### Chao richness
**ANOVA**
```{r}
metaAave %>%
  filter(Type == "L" & AgeGroup == "2yr") %>% 
  aov(chao ~ Effic+MPE, .) %>% 
  summary()
```

### Rumen solids
#### Shannon's diversity
**ANOVA**
```{r}
metaAave %>%
  filter(Type == "S" & AgeGroup == "2yr") %>% 
  aov(shannon ~ Effic+MPE, .) %>% 
  summary()
```

#### Chao richness
**ANOVA**
```{r}
metaAave %>%
  filter(Type == "S" & AgeGroup == "2yr") %>% 
  aov(chao ~ Effic+MPE, .) %>% 
  summary()
```

## Bacteria
### Fecal
#### Shannon's diversity
**ANOVA**
```{r}
metaBave %>%
  filter(Type == "F" & AgeGroup == "2yr") %>% 
  aov(shannon ~ Effic+MPE, .) %>% 
  summary()
```

#### Chao richness
**ANOVA**
```{r}
metaBave %>%
  filter(Type == "F" & AgeGroup == "2yr") %>% 
  aov(chao ~ Effic+MPE, .) %>% 
  summary()
```

### Rumen liquids
#### Shannon's diversity
**ANOVA**
```{r}
metaBave %>%
  filter(Type == "L" & AgeGroup == "2yr") %>% 
  aov(shannon ~ Effic+MPE, .) %>% 
  summary()
```

#### Chao richness
**ANOVA**
```{r}
metaBave %>%
  filter(Type == "L" & AgeGroup == "2yr") %>% 
  aov(chao ~ Effic+MPE, .) %>% 
  summary()
```

### Rumen solids
#### Shannon's diversity
**ANOVA**
```{r}
metaBave %>%
  filter(Type == "S" & AgeGroup == "2yr") %>% 
  aov(shannon ~ Effic+MPE, .) %>% 
  summary()
```

#### Chao richness
**ANOVA**
```{r}
metaBave %>%
  filter(Type == "S" & AgeGroup == "2yr") %>% 
  aov(chao ~ Effic+MPE, .) %>% 
  summary()
```

## Fungi
### Fecal
#### Shannon's diversity
**ANOVA**
```{r}
metaFave %>%
  filter(Type == "F" & AgeGroup == "2yr") %>% 
  aov(shannon ~ Effic+MPE, .) %>% 
  summary()
```

#### Chao richness
**ANOVA**
```{r}
metaFave %>%
  filter(Type == "F" & AgeGroup == "2yr") %>% 
  aov(chao ~ Effic+MPE, .) %>% 
  summary()
```

### Rumen liquids
#### Shannon's diversity
**ANOVA**
```{r}
metaFave %>%
  filter(Type == "L" & AgeGroup == "2yr") %>% 
  aov(shannon ~ Effic+MPE, .) %>% 
  summary()
```

#### Chao richness
**ANOVA**
```{r}
metaFave %>%
  filter(Type == "L" & AgeGroup == "2yr") %>% 
  aov(chao ~ Effic+MPE, .) %>% 
  summary()
```

### Rumen solids
#### Shannon's diversity
**ANOVA**
```{r}
metaFave %>%
  filter(Type == "S" & AgeGroup == "2yr") %>% 
  aov(shannon ~ Effic+MPE, .) %>% 
  summary()
```

#### Chao richness
**ANOVA**
```{r}
metaFave %>%
  filter(Type == "S" & AgeGroup == "2yr") %>% 
  aov(chao ~ Effic+MPE, .) %>% 
  summary()
```
