---
title: "Beta-diversity analyses"
subtitle: "GS01 Calf diets"
author: "Kim Dill-McFarland"
date: "version `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_float:
      collapsed: no
editor_options: 
  chunk_output_type: console
---
<script>
$(document).ready(function() {
  $items = $('div#TOC li');
  $items.each(function(idx) {
    num_ul = $(this).parentsUntil('#TOC').length;
    $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
  });
});
</script>

# Load data
Created in `1_data_manip.Rmd`
```{r}
set.seed(8325)
load("GS01.data.RData")
```

# Packages
```{r Load packages}
#Beta-diversity analyses and plots
library(vegan)
#Data manipulation
library(tidyverse)
```

# Functions
```{r}
source("functions/Pairwise_adonis2.R")
source("functions/pairwise_permdisp.R")
```

# Average beta by age
## Archaea
**Bray-Curtis**
```{r}
BC.temp = OTUA %>%
  # Remove non-numeric labels
  select(-Sample) %>% 
  vegdist(method="bray") %>% 
  # Convert to data frame
  as.matrix() %>% 
  as.data.frame() %>% 
  # Label columns and rows by sample name
  mutate(Sample=OTUA$Sample) %>%
  rename_if(is.numeric, ~ OTUA$Sample) %>% 
  # Add metadata
  full_join(metaA[,c("Sample","Animal","AgeGroup","Type")], by="Sample") %>% 
  # Reformat to long
  gather(-Sample, -Animal, -AgeGroup, -Type, key="Sample2", value="BC") %>% 
  # Remove self comparisons
  dplyr::filter(BC > 0) %>% 
  # Calculate mean and stdev
  group_by(AgeGroup, Type) %>% 
  summarize(mean=mean(BC, na.rm=TRUE), sd=sd(BC, na.rm=TRUE)) %>% 
  mutate(amplicon="Arc", metric="BC")
```

**Jaccard**
```{r}
J.temp = OTUA %>%
  # Remove non-numeric labels
  select(-Sample) %>% 
  vegdist(method="jaccard") %>% 
  # Convert to data frame
  as.matrix() %>% 
  as.data.frame() %>% 
  # Label columns and rows by sample name
  mutate(Sample=OTUA$Sample) %>%
  rename_if(is.numeric, ~ OTUA$Sample) %>% 
  # Add metadata
  full_join(metaA[,c("Sample","Animal","AgeGroup","Type")], by="Sample") %>% 
  # Reformat to long
  gather(-Sample, -Animal, -AgeGroup, -Type, key="Sample2", value="J") %>% 
  # Remove self comparisons
  dplyr::filter(J > 0) %>% 
  # Calculate mean and stdev
  group_by(AgeGroup, Type) %>% 
  summarize(mean=mean(J, na.rm=TRUE), sd=sd(J, na.rm=TRUE)) %>% 
  mutate(amplicon="Arc", metric="J")
```

**Combine**
```{r eval=FALSE}
write.table(rbind(BC.temp, J.temp), "results/beta_div/A.beta.txt", row.names=FALSE, sep="\t")
```

## Bacteria
**Bray-Curtis**
```{r}
BC.temp = OTUB %>%
  # Remove non-numeric labels
  select(-Sample) %>% 
  vegdist(method="bray") %>% 
  # Convert to data frame
  as.matrix() %>% 
  as.data.frame() %>% 
  # Label columns and rows by sample name
  mutate(Sample=OTUB$Sample) %>%
  rename_if(is.numeric, ~ OTUB$Sample) %>% 
  # Add metadata
  full_join(metaB[,c("Sample","Animal","AgeGroup","Type")], by="Sample") %>% 
  # Reformat to long
  gather(-Sample, -Animal, -AgeGroup, -Type, key="Sample2", value="BC") %>% 
  # Remove self comparisons
  dplyr::filter(BC > 0) %>% 
  # Calculate mean and stdev
  group_by(AgeGroup, Type) %>% 
  summarize(mean=mean(BC, na.rm=TRUE), sd=sd(BC, na.rm=TRUE)) %>% 
  mutate(amplicon="Bac", metric="BC")
```

**Jaccard**
```{r}
J.temp = OTUB %>%
  # Remove non-numeric labels
  select(-Sample) %>% 
  vegdist(method="jaccard") %>% 
  # Convert to data frame
  as.matrix() %>% 
  as.data.frame() %>% 
  # Label columns and rows by sample name
  mutate(Sample=OTUB$Sample) %>%
  rename_if(is.numeric, ~ OTUB$Sample) %>% 
  # Add metadata
  full_join(metaB[,c("Sample","Animal","AgeGroup","Type")], by="Sample") %>% 
  # Reformat to long
  gather(-Sample, -Animal, -AgeGroup, -Type, key="Sample2", value="J") %>% 
  # Remove self comparisons
  dplyr::filter(J > 0) %>% 
  # Calculate mean and stdev
  group_by(AgeGroup, Type) %>% 
  summarize(mean=mean(J, na.rm=TRUE), sd=sd(J, na.rm=TRUE)) %>% 
  mutate(amplicon="Bac", metric="J")
```

**Combine**
```{r eval=FALSE}
write.table(rbind(BC.temp, J.temp), "results/beta_div/B.beta.txt", row.names=FALSE, sep="\t")
```

## Fungi
**Bray-Curtis**
```{r}
BC.temp = OTUF %>%
  # Remove non-numeric labels
  select(-Sample) %>% 
  vegdist(method="bray") %>% 
  # Convert to data frame
  as.matrix() %>% 
  as.data.frame() %>% 
  # Label columns and rows by sample name
  mutate(Sample=OTUF$Sample) %>%
  rename_if(is.numeric, ~ OTUF$Sample) %>% 
  # Add metadata
  full_join(metaF[,c("Sample","Animal","AgeGroup","Type")], by="Sample") %>% 
  # Reformat to long
  gather(-Sample, -Animal, -AgeGroup, -Type, key="Sample2", value="BC") %>% 
  # Remove self comparisons
  dplyr::filter(BC > 0) %>% 
  # Calculate mean and stdev
  group_by(AgeGroup, Type) %>% 
  summarize(mean=mean(BC, na.rm=TRUE), sd=sd(BC, na.rm=TRUE)) %>% 
  mutate(amplicon="Fun", metric="BC")
```

**Jaccard**
```{r}
J.temp = OTUF %>%
  # Remove non-numeric labels
  select(-Sample) %>% 
  vegdist(method="jaccard") %>% 
  # Convert to data frame
  as.matrix() %>% 
  as.data.frame() %>% 
  # Label columns and rows by sample name
  mutate(Sample=OTUF$Sample) %>%
  rename_if(is.numeric, ~ OTUF$Sample) %>% 
  # Add metadata
  full_join(metaF[,c("Sample","Animal","AgeGroup","Type")], by="Sample") %>% 
  # Reformat to long
  gather(-Sample, -Animal, -AgeGroup, -Type, key="Sample2", value="J") %>% 
  # Remove self comparisons
  dplyr::filter(J > 0) %>% 
  # Calculate mean and stdev
  group_by(AgeGroup, Type) %>% 
  summarize(mean=mean(J, na.rm=TRUE), sd=sd(J, na.rm=TRUE)) %>% 
  mutate(amplicon="Fun", metric="J")
```

**Combine**
```{r eval=FALSE}
write.table(rbind(BC.temp, J.temp), "results/beta_div/F.beta.txt", row.names=FALSE, sep="\t")
```

## Combine all
```{r message=FALSE}
beta.ave = read_tsv("results/beta_div/A.beta.txt") %>%
  rbind(read_tsv("results/beta_div/B.beta.txt")) %>% 
  rbind(read_tsv("results/beta_div/F.beta.txt")) %>%
  gather(mean:sd, key=stat, value=value) %>% 
  mutate(group=paste(AgeGroup, stat)) %>% 
  select(-stat,-AgeGroup) %>% 
  group_by(Type, amplicon, metric) %>% 
  spread(key=group, value=value)
```

## Calf vs. Cow
Average calf (2, 4, 8w) and cow (1, 2yr) sample groups
```{r message=FALSE}
read_tsv("results/beta_div/A.beta.txt") %>%
  rbind(read_tsv("results/beta_div/B.beta.txt")) %>% 
  rbind(read_tsv("results/beta_div/F.beta.txt")) %>%
  mutate(group=ifelse(AgeGroup %in% c("2w","4w","8w"), "calf", "cow")) %>% 
  group_by(Type, amplicon, metric, group) %>% 
  summarize(mean2=mean(mean, na.rm=TRUE), sd2=sd(sd, na.rm=TRUE)) %>% 
  gather(mean2:sd2, key=stat, value=value) %>% 
  mutate(group2=paste(group, stat)) %>% 
  select(-group,-stat) %>% 
  group_by(Type, amplicon, metric) %>% 
  spread(key=group2, value=value) %>% 
  full_join(beta.ave, by=c("Type","amplicon","metric")) %>% 
  write.table("results/beta_div/all.beta.txt", row.names=FALSE, sep="\t")
```

# Model 1: Diet and age
Diet * AgeGroup

* Permutational ANOVA (PERMANOVA) followed by pairwise PERMANOVA with FDR correction (Benjamini-Hochberg method) if significant
* Permutation test of multivariate homogeneity of group dispersions (PERMDISP) followed by pairwise TukeyHSD tests if significant

## Archaea
### Fecal
Subset data to fecal samples
```{r}
meta.temp = metaA %>%
  filter(Type == "F") %>% 
  arrange(Sample)
```

```{r}
OTU.temp = metaA %>%
  select(Sample, Type, Diet, AgeGroup, DietAge) %>% 
  filter(Type == "F") %>% 
  left_join(OTUA, by="Sample") %>% 
  arrange(Sample)
```

#### PERMANOVA
Bray-Curtis (structure)
```{r}
adonis(OTU.temp[, grepl("Otu", colnames(OTU.temp))] ~ Diet*AgeGroup, data=meta.temp, method="bray", permutations=1000)
```

```{r eval=FALSE}
pairwise.adonis(OTU.temp, meta.temp, c("Diet", "AgeGroup", "DietAge"), "bray", "fdr", "results/adonis_OTU/A.F.OTU.bray")
```

Jaccard (composition)
```{r}
adonis(OTU.temp[, grepl("Otu", colnames(OTU.temp))] ~ Diet*AgeGroup, data=meta.temp, method="jaccard", permutations=1000)
```

```{r eval=FALSE}
pairwise.adonis(OTU.temp, meta.temp, c("Diet", "AgeGroup", "DietAge"), "jaccard", "fdr", "results/adonis_OTU/A.F.OTU.jac")
```

#### PERMDISP
Bray-Curtis
```{r eval=FALSE}
pairwise.permdisp(OTU.temp, meta.temp, c("Diet", "AgeGroup", "DietAge"), "bray", "results/permdisp_OTU/A.F.OTU.bray")
```

### Rumen liquids
Subset data to liquid samples
```{r}
OTU.temp = metaA %>%
  select(Sample, Type, Diet, AgeGroup, DietAge) %>% 
  filter(Type == "L") %>% 
  left_join(OTUA, by="Sample") %>% 
  arrange(Sample)

meta.temp = metaA %>%
  filter(Type == "L") %>% 
  arrange(Sample)
```

#### PERMANOVA
Bray-Curtis (structure)
```{r}
adonis(OTU.temp[, grepl("Otu", colnames(OTU.temp))] ~ Diet*AgeGroup, data=meta.temp, method="bray", permutations=1000)
```

```{r eval=FALSE}
pairwise.adonis(OTU.temp, meta.temp, c("Diet", "AgeGroup", "DietAge"), "bray", "fdr", "results/adonis_OTU/A.L.OTU.bray")
```

Jaccard (composition)
```{r}
adonis(OTU.temp[, grepl("Otu", colnames(OTU.temp))] ~ Diet*AgeGroup, data=meta.temp, method="jaccard", permutations=1000)
```

```{r eval=FALSE}
pairwise.adonis(OTU.temp, meta.temp, c("Diet", "AgeGroup", "DietAge"), "jaccard", "fdr", "results/adonis_OTU/A.L.OTU.jac")
```

#### PERMDISP
Bray-Curtis
```{r eval=FALSE}
pairwise.permdisp(OTU.temp, meta.temp, c("Diet", "AgeGroup", "DietAge"), "bray", "results/permdisp_OTU/A.L.OTU.bray")
```

### Rumen solids
Subset data to solid samples
```{r}
OTU.temp = metaA %>%
  select(Sample, Type, Diet, AgeGroup, DietAge) %>% 
  filter(Type == "S") %>% 
  left_join(OTUA, by="Sample") %>% 
  arrange(Sample)

meta.temp = metaA %>%
  filter(Type == "S") %>% 
  arrange(Sample)
```

#### PERMANOVA
Bray-Curtis (structure)
```{r}
adonis(OTU.temp[, grepl("Otu", colnames(OTU.temp))] ~ Diet*AgeGroup, data=meta.temp, method="bray", permutations=1000)
```

```{r eval=FALSE}
pairwise.adonis(OTU.temp, meta.temp, c("Diet", "AgeGroup", "DietAge"), "bray", "fdr", "results/adonis_OTU/A.S.OTU.bray")
```

Jaccard (composition)
```{r}
adonis(OTU.temp[, grepl("Otu", colnames(OTU.temp))] ~ Diet*AgeGroup, data=meta.temp, method="jaccard", permutations=1000)
```

```{r eval=FALSE}
pairwise.adonis(OTU.temp, meta.temp, c("Diet", "AgeGroup", "DietAge"), "jaccard", "fdr", "results/adonis_OTU/A.S.OTU.jac")
```

#### PERMDISP
Bray-Curtis
```{r eval=FALSE}
pairwise.permdisp(OTU.temp, meta.temp, c("Diet", "AgeGroup", "DietAge"), "bray", "results/permdisp_OTU/A.S.OTU.bray")
```

## Bacteria
### Fecal
Subset data to fecal samples
```{r}
OTU.temp = metaB %>%
  select(Sample, Type, Diet, AgeGroup, DietAge) %>% 
  filter(Type == "F") %>% 
  left_join(OTUB, by="Sample") %>% 
  arrange(Sample)

meta.temp = metaB %>%
  filter(Type == "F") %>% 
  arrange(Sample)
```

#### PERMANOVA
Bray-Curtis (structure)
```{r}
adonis(OTU.temp[, grepl("Otu", colnames(OTU.temp))] ~ Diet*AgeGroup, data=meta.temp, method="bray", permutations=1000)
```

```{r eval=FALSE}
pairwise.adonis(OTU.temp, meta.temp, c("Diet", "AgeGroup", "DietAge"), "bray", "fdr", "results/adonis_OTU/B.F.OTU.bray")
```

Jaccard (composition)
```{r}
adonis(OTU.temp[, grepl("Otu", colnames(OTU.temp))] ~ Diet*AgeGroup, data=meta.temp, method="jaccard", permutations=1000)
```

```{r eval=FALSE}
pairwise.adonis(OTU.temp, meta.temp, c("Diet", "AgeGroup", "DietAge"), "jaccard", "fdr", "results/adonis_OTU/B.F.OTU.jac")
```

#### PERMDISP
Bray-Curtis
```{r eval=FALSE}
pairwise.permdisp(OTU.temp, meta.temp, c("Diet", "AgeGroup", "DietAge"), "bray", "results/permdisp_OTU/B.F.OTU.bray")
```

### Rumen liquids
Subset data to liquid samples
```{r}
OTU.temp = metaB %>%
  select(Sample, Type, Diet, AgeGroup, DietAge) %>% 
  filter(Type == "L") %>% 
  left_join(OTUB, by="Sample") %>% 
  arrange(Sample)

meta.temp = metaB %>%
  filter(Type == "L") %>% 
  arrange(Sample)
```

#### PERMANOVA
Bray-Curtis (structure)
```{r}
adonis(OTU.temp[, grepl("Otu", colnames(OTU.temp))] ~ Diet*AgeGroup, data=meta.temp, method="bray", permutations=1000)
```

```{r eval=FALSE}
pairwise.adonis(OTU.temp, meta.temp, c("Diet", "AgeGroup", "DietAge"), "bray", "fdr", "results/adonis_OTU/B.L.OTU.bray")
```

Jaccard (composition)
```{r}
adonis(OTU.temp[, grepl("Otu", colnames(OTU.temp))] ~ Diet*AgeGroup, data=meta.temp, method="jaccard", permutations=1000)
```

```{r eval=FALSE}
pairwise.adonis(OTU.temp, meta.temp, c("Diet", "AgeGroup", "DietAge"), "jaccard", "fdr", "results/adonis_OTU/B.L.OTU.jac")
```

#### PERMDISP
Bray-Curtis
```{r eval=FALSE}
pairwise.permdisp(OTU.temp, meta.temp, c("Diet", "AgeGroup", "DietAge"), "bray", "results/permdisp_OTU/B.L.OTU.bray")
```

### Rumen solids
Subset data to solid samples
```{r}
OTU.temp = metaB %>%
  select(Sample, Type, Diet, AgeGroup, DietAge) %>% 
  filter(Type == "S") %>% 
  left_join(OTUB, by="Sample") %>% 
  arrange(Sample)

meta.temp = metaB %>%
  filter(Type == "S") %>% 
  arrange(Sample)
```

#### PERMANOVA
Bray-Curtis (structure)
```{r}
adonis(OTU.temp[, grepl("Otu", colnames(OTU.temp))] ~ Diet*AgeGroup, data=meta.temp, method="bray", permutations=1000)
```

```{r eval=FALSE}
pairwise.adonis(OTU.temp, meta.temp, c("Diet", "AgeGroup", "DietAge"), "bray", "fdr", "results/adonis_OTU/B.S.OTU.bray")
```

Jaccard (composition)
```{r}
adonis(OTU.temp[, grepl("Otu", colnames(OTU.temp))] ~ Diet*AgeGroup, data=meta.temp, method="jaccard", permutations=1000)
```

```{r eval=FALSE}
pairwise.adonis(OTU.temp, meta.temp, c("Diet", "AgeGroup", "DietAge"), "jaccard", "fdr", "results/adonis_OTU/B.S.OTU.jac")
```

#### PERMDISP
Bray-Curtis
```{r eval=FALSE}
pairwise.permdisp(OTU.temp, meta.temp, c("Diet", "AgeGroup", "DietAge"), "bray", "results/permdisp_OTU/B.S.OTU.bray")
```

## Fungi
### Fecal
Subset data to fecal samples
```{r}
OTU.temp = metaF %>%
  select(Sample, Type, Diet, AgeGroup, DietAge) %>% 
  filter(Type == "F") %>% 
  left_join(OTUF, by="Sample") %>% 
  arrange(Sample)

meta.temp = metaF %>%
  filter(Type == "F")%>% 
  arrange(Sample)
```

#### PERMANOVA
Bray-Curtis (structure)
```{r}
adonis(OTU.temp[, grepl("Otu", colnames(OTU.temp))] ~ Diet*AgeGroup, data=meta.temp, method="bray", permutations=1000)
```

```{r eval=FALSE}
pairwise.adonis(OTU.temp, meta.temp, c("Diet", "AgeGroup", "DietAge"), "bray", "fdr", "results/adonis_OTU/F.F.OTU.bray")
```

Jaccard (composition)
```{r}
adonis(OTU.temp[, grepl("Otu", colnames(OTU.temp))] ~ Diet*AgeGroup, data=meta.temp, method="jaccard", permutations=1000)
```

```{r eval=FALSE}
pairwise.adonis(OTU.temp, meta.temp, c("Diet", "AgeGroup", "DietAge"), "jaccard", "fdr", "results/adonis_OTU/F.F.OTU.jac")
```

#### PERMDISP
Bray-Curtis
```{r eval=FALSE}
pairwise.permdisp(OTU.temp, meta.temp, c("Diet", "AgeGroup", "DietAge"), "bray", "results/permdisp_OTU/F.F.OTU.bray")
```

### Rumen liquids
Subset data to liquid samples
```{r}
OTU.temp = metaF %>%
  select(Sample, Type, Diet, AgeGroup, DietAge) %>% 
  filter(Type == "L") %>% 
  left_join(OTUF, by="Sample") %>% 
  arrange(Sample)

meta.temp = metaF %>%
  filter(Type == "L") %>% 
  arrange(Sample)
```

#### PERMANOVA
Bray-Curtis (structure)
```{r}
adonis(OTU.temp[, grepl("Otu", colnames(OTU.temp))] ~ Diet*AgeGroup, data=meta.temp, method="bray", permutations=1000)
```

```{r eval=FALSE}
pairwise.adonis(OTU.temp, meta.temp, c("Diet", "AgeGroup", "DietAge"), "bray", "fdr", "results/adonis_OTU/F.L.OTU.bray")
```

Jaccard (composition)
```{r}
adonis(OTU.temp[, grepl("Otu", colnames(OTU.temp))] ~ Diet*AgeGroup, data=meta.temp, method="jaccard", permutations=1000)
```

```{r eval=FALSE}
pairwise.adonis(OTU.temp, meta.temp, c("Diet", "AgeGroup", "DietAge"), "jaccard", "fdr", "results/adonis_OTU/F.L.OTU.jac")
```

#### PERMDISP
Bray-Curtis
```{r eval=FALSE}
pairwise.permdisp(OTU.temp, meta.temp, c("Diet", "AgeGroup", "DietAge"), "bray", "results/permdisp_OTU/F.L.OTU.bray")
```

### Rumen solids
Subset data to solid samples
```{r}
OTU.temp = metaF %>%
  select(Sample, Type, Diet, AgeGroup, DietAge) %>% 
  filter(Type == "S") %>% 
  left_join(OTUF, by="Sample") %>% 
  arrange(Sample)

meta.temp = metaF %>%
  filter(Type == "S") %>% 
  arrange(Sample)
```

#### PERMANOVA
Bray-Curtis (structure)
```{r}
adonis(OTU.temp[, grepl("Otu", colnames(OTU.temp))] ~ Diet*AgeGroup, data=meta.temp, method="bray", permutations=1000)
```

```{r eval=FALSE}
pairwise.adonis(OTU.temp, meta.temp, c("Diet", "AgeGroup", "DietAge"), "bray", "fdr", "results/adonis_OTU/F.S.OTU.bray")
```

Jaccard (composition)
```{r}
adonis(OTU.temp[, grepl("Otu", colnames(OTU.temp))] ~ Diet*AgeGroup, data=meta.temp, method="jaccard", permutations=1000)
```

```{r eval=FALSE}
pairwise.adonis(OTU.temp, meta.temp, c("Diet", "AgeGroup", "DietAge"), "jaccard", "fdr", "results/adonis_OTU/F.S.OTU.jac")
```

#### PERMDISP
Bray-Curtis
```{r eval=FALSE}
pairwise.permdisp(OTU.temp, meta.temp, c("Diet", "AgeGroup", "DietAge"), "bray", "results/permdisp_OTU/F.S.OTU.bray")
```

## Combine PERMANOVA results
```{r}
{# Set working directory for results files
  require(plyr)
  setwd("results/adonis_OTU/")
# List all Kruskal results files
file_list = list.files()

  # Read in all files and merge into 1 data frame
for(file in file_list){
  # if the merged dataset doesn't exist, create it
  if (!exists("adonis_OTU_results")){
    # Read in 1st dataset
    adonis_OTU_results = read.table(file, header=TRUE, row.names=NULL, sep="\t")
  # Add file name to data column to identify origin
    adonis_OTU_results$origin = file
  }
   
  # if the merged dataset does exist, append to it
  if (exists("adonis_OTU_results")){
    # Read in additional dataset
    temp_dataset =  read.table(file, header=TRUE, row.names=NULL, sep="\t")
    # Add file name to data column to identify origin
    temp_dataset$origin = file
    # Merge datasets
    adonis_OTU_results<-rbind.fill(adonis_OTU_results, temp_dataset)
    # Remove original additional dataset
    rm(temp_dataset)
  }
}
# Return working directory to main project directory
setwd("../..")
}
```

```{r}
adonis_OTU_results %>% 
  dplyr::filter(Comparison %in%
                c("A_B", "A_C", "B_C",
                  "B_A", "C_A", "C_B",
                  
                  "2w_4w", "4w_8w",  "8w_1yr","1yr_2yr",
                  "4w_2w", "8w_4w",  "1yr_8w","2yr_1yr",
                  
                  "A2w_B2w", "A2w_C2w", "B2w_C2w", 
                  "B2w_A2w", "C2w_A2w", "C2w_B2w",
                  "A4w_B4w", "A4w_C4w", "B4w_C4w", 
                  "B4w_A4w", "C4w_A4w", "C4w_B4w", 
                  "A8w_B8w", "A8w_C8w", "B8w_C8w", 
                  "B8w_A8w", "C8w_A8w", "C8w_B8w",
                  "A1yr_B1yr", "A1yr_C1yr", "B1yr_C1yr", 
                  "B1yr_A1yr", "C1yr_A1yr", "C1yr_B1yr",
                  "A2yr_B2yr", "A2yr_C2yr", "B2yr_C2yr",
                  "B2yr_A2yr", "C2yr_A2yr", "C2yr_B2yr")) %>% 
  select(-Rsq, -pval) %>% 
  distinct() %>% 
  spread(key=Comparison,  value=p.adj) %>% 
  separate(origin, c("amplicon", "Type", "variable", "beta", "test", "filetype"), by=c(".","_")) %>% 
  write_tsv("results/adonis_OTU_results.txt", col_names=TRUE)
```

## Combine PERMDISP results
```{r}
{# Set working directory for results files
  require(plyr)
  setwd("results/permdisp_OTU/")
# List all Kruskal results files
file_list = list.files()

  # Read in all files and merge into 1 data frame
for(file in file_list){
  # if the merged dataset doesn't exist, create it
  if (!exists("permdisp_OTU_results")){
    # Read in 1st dataset
    permdisp_OTU_results = read.table(file, header=TRUE, row.names=NULL, sep="\t")
  # Add file name to data column to identify origin
    permdisp_OTU_results$origin = file
  }
   
  # if the merged dataset does exist, append to it
  if (exists("permdisp_OTU_results")){
    # Read in additional dataset
    temp_dataset =  read.table(file, header=TRUE, row.names=NULL, sep="\t")
    # Add file name to data column to identify origin
    temp_dataset$origin = file
    # Merge datasets
    permdisp_OTU_results<-rbind.fill(permdisp_OTU_results, temp_dataset)
    # Remove original additional dataset
    rm(temp_dataset)
  }
}
# Return working directory to main project directory
setwd("../..")
}
```

```{r}
permdisp_OTU_results %>% 
  dplyr::filter(label %in%
                c("A-B", "A-C", "B-C",
                  "B-A", "C-A", "C-B",
                  
                  "2w-4w", "4w-8w",  "8w-1yr","1yr-2yr",
                  "4w-2w", "8w-4w",  "1yr-8w","2yr-1yr",
                  
                  "A2w-B2w", "A2w-C2w", "B2w-C2w", 
                  "B2w-A2w", "C2w-A2w", "C2w-B2w",
                  "A4w-B4w", "A4w-C4w", "B4w-C4w", 
                  "B4w-A4w", "C4w-A4w", "C4w-B4w", 
                  "A8w-B8w", "A8w-C8w", "B8w-C8w", 
                  "B8w-A8w", "C8w-A8w", "C8w-B8w",
                  "A1yr-B1yr", "A1yr-C1yr", "B1yr-C1yr", 
                  "B1yr-A1yr", "C1yr-A1yr", "C1yr-B1yr",
                  "A2yr-B2yr", "A2yr-C2yr", "B2yr-C2yr",
                  "B2yr-A2yr", "C2yr-A2yr", "C2yr-B2yr", "1")) %>% 
  distinct() %>% 
  spread(key=label, value=pval) %>% 
  separate(origin, c("amplicon", "Type", "variable", "beta", "test", "filetype"), by=c(".","_")) %>% 
  write_tsv("results/permdisp_OTU_results.txt", col_names=TRUE)
```

# Model 2: Milk production efficiency
2yr samples only, averaged 3 consecutive rumen samples to avoid animal effects  
Effic + MPE

* Permutational ANOVA (PERMANOVA) followed by pairwise PERMANOVA with FDR correction (Benjamini-Hochberg method) if significant

## Archaea
### Fecal
Subset data to 2yr fecal samples
```{r}
## Metadata
meta.temp = metaAave %>%
  filter(Type == "F" & AgeGroup == "2yr")
```

```{r}
OTU.temp = metaAave %>%
  # Combine with metadata
  select(Sample, AgeGroup, Type) %>% 
  right_join(OTUAave, by="Sample") %>%
  # Filter to only 2 yr fecal samples
  filter(Type == "F" & AgeGroup == "2yr") %>% 
  # Remove metadata variables
  select(-Sample, -AgeGroup, -Type)
```

**PERMANOVA**  
Bray-Curtis (structure)
```{r}
adonis(OTU.temp ~ Effic+MPE, data=meta.temp, method="bray", permutations=1000)
```

Jaccard (composition)
```{r}
adonis(OTU.temp ~ Effic+MPE, data=meta.temp, method="jaccard", permutations=1000)
```

### Rumen liquids
Subset data to 2yr rumen liquid samples
```{r}
## Metadata
meta.temp = metaAave %>%
  # Efficieny 
  filter(Type == "L" & AgeGroup == "2yr")
```

```{r}
OTU.temp = metaAave %>%
  # Combine with metadata
  select(Sample, AgeGroup, Type) %>% 
  right_join(OTUAave, by="Sample") %>%
  # Filter to only 2 yr fecal samples
  filter(Type == "L" & AgeGroup == "2yr") %>% 
  # Remove metadata variables
  select(-Sample, -AgeGroup, -Type)
```

**PERMANOVA**  
Bray-Curtis (structure)
```{r}
adonis(OTU.temp ~ Effic+MPE, data=meta.temp, method="bray", permutations=1000)
```

Jaccard (composition)
```{r}
adonis(OTU.temp ~ Effic+MPE, data=meta.temp, method="jaccard", permutations=1000)
```

### Rumen solids
Subset data to 2yr rumen solid samples
```{r}
## Metadata
meta.temp = metaAave %>%
  # Efficieny 
  filter(Type == "S" & AgeGroup == "2yr")
```

```{r}
OTU.temp = metaAave %>%
  # Combine with metadata
  select(Sample, AgeGroup, Type) %>% 
  right_join(OTUAave, by="Sample") %>%
  # Filter to only 2 yr fecal samples
  filter(Type == "S" & AgeGroup == "2yr") %>% 
  # Remove metadata variables
  select(-Sample, -AgeGroup, -Type)
```

**PERMANOVA**  
Bray-Curtis (structure)
```{r}
adonis(OTU.temp ~ Effic+MPE, data=meta.temp, method="bray", permutations=1000)
```

Jaccard (composition)
```{r}
adonis(OTU.temp ~ Effic+MPE, data=meta.temp, method="jaccard", permutations=1000)
```

## Bacteria
### Fecal
Subset data to 2yr fecal samples
```{r}
## Metadata
meta.temp = metaBave %>%
  filter(Type == "F" & AgeGroup == "2yr")
```

```{r}
OTU.temp = metaBave %>%
  # Combine with metadata
  select(Sample, AgeGroup, Type) %>% 
  right_join(OTUBave, by="Sample") %>%
  # Filter to only 2 yr fecal samples
  filter(Type == "F" & AgeGroup == "2yr") %>% 
  # Remove metadata variables
  select(-Sample, -AgeGroup, -Type)
```

**PERMANOVA**  
Bray-Curtis (structure)
```{r}
adonis(OTU.temp ~ Effic+MPE, data=meta.temp, method="bray", permutations=1000)
```

Jaccard (composition)
```{r}
adonis(OTU.temp ~ Effic+MPE, data=meta.temp, method="jaccard", permutations=1000)
```

### Rumen liquids
Subset data to 2yr rumen liquid samples
```{r}
## Metadata
meta.temp = metaBave %>%
  # Efficieny 
  filter(Type == "L" & AgeGroup == "2yr")
```

```{r}
OTU.temp = metaBave %>%
  # Combine with metadata
  select(Sample, AgeGroup, Type) %>% 
  right_join(OTUBave, by="Sample") %>%
  # Filter to only 2 yr fecal samples
  filter(Type == "L" & AgeGroup == "2yr") %>% 
  # Remove metadata variables
  select(-Sample, -AgeGroup, -Type)
```

**PERMANOVA**  
Bray-Curtis (structure)
```{r}
adonis(OTU.temp ~ Effic+MPE, data=meta.temp, method="bray", permutations=1000)
```

Jaccard (composition)
```{r}
adonis(OTU.temp ~ Effic+MPE, data=meta.temp, method="jaccard", permutations=1000)
```

### Rumen solids
Subset data to 2yr rumen solid samples
```{r}
## Metadata
meta.temp = metaBave %>%
  # Efficieny 
  filter(Type == "S" & AgeGroup == "2yr")
```

```{r}
OTU.temp = metaBave %>%
  # Combine with metadata
  select(Sample, AgeGroup, Type) %>% 
  right_join(OTUBave, by="Sample") %>%
  # Filter to only 2 yr fecal samples
  filter(Type == "S" & AgeGroup == "2yr") %>% 
  # Remove metadata variables
  select(-Sample, -AgeGroup, -Type)
```

**PERMANOVA**  
Bray-Curtis (structure)
```{r}
adonis(OTU.temp ~ Effic+MPE, data=meta.temp, method="bray", permutations=1000)
```

Jaccard (composition)
```{r}
adonis(OTU.temp ~ Effic+MPE, data=meta.temp, method="jaccard", permutations=1000)
```

## Fungi
### Fecal
Subset data to 2yr fecal samples
```{r}
## Metadata
meta.temp = metaFave %>%
  filter(Type == "F" & AgeGroup == "2yr")
```

```{r}
OTU.temp = metaFave %>%
  # Combine with metadata
  select(Sample, AgeGroup, Type) %>% 
  right_join(OTUFave, by="Sample") %>%
  # Filter to only 2 yr fecal samples
  filter(Type == "F" & AgeGroup == "2yr") %>% 
  # Remove metadata variables
  select(-Sample, -AgeGroup, -Type)
```

**PERMANOVA**  
Bray-Curtis (structure)
```{r}
adonis(OTU.temp ~ Effic+MPE, data=meta.temp, method="bray", permutations=1000)
```

Jaccard (composition)
```{r}
adonis(OTU.temp ~ Effic+MPE, data=meta.temp, method="jaccard", permutations=1000)
```

### Rumen liquids
Subset data to 2yr rumen liquid samples
```{r}
## Metadata
meta.temp = metaFave %>%
  # Efficieny 
  filter(Type == "L" & AgeGroup == "2yr")
```

```{r}
OTU.temp = metaFave %>%
  # Combine with metadata
  select(Sample, AgeGroup, Type) %>% 
  right_join(OTUFave, by="Sample") %>%
  # Filter to only 2 yr fecal samples
  filter(Type == "L" & AgeGroup == "2yr") %>% 
  # Remove metadata variables
  select(-Sample, -AgeGroup, -Type)
```

**PERMANOVA**  
Bray-Curtis (structure)
```{r}
adonis(OTU.temp ~ Effic+MPE, data=meta.temp, method="bray", permutations=1000)
```

Jaccard (composition)
```{r}
adonis(OTU.temp ~ Effic+MPE, data=meta.temp, method="jaccard", permutations=1000)
```

### Rumen solids
Subset data to 2yr rumen solid samples
```{r}
## Metadata
meta.temp = metaFave %>%
  # Efficieny 
  filter(Type == "S" & AgeGroup == "2yr")
```

```{r}
OTU.temp = metaFave %>%
  # Combine with metadata
  select(Sample, AgeGroup, Type) %>% 
  right_join(OTUFave, by="Sample") %>%
  # Filter to only 2 yr fecal samples
  filter(Type == "S" & AgeGroup == "2yr") %>% 
  # Remove metadata variables
  select(-Sample, -AgeGroup, -Type)
```

**PERMANOVA**  
Bray-Curtis (structure)
```{r}
adonis(OTU.temp ~ Effic+MPE, data=meta.temp, method="bray", permutations=1000)
```

Jaccard (composition)
```{r}
adonis(OTU.temp ~ Effic+MPE, data=meta.temp, method="jaccard", permutations=1000)
```

# Model 3: Calf scours
2- and 4-week samples only, stratify by AgeGroup   ScourSample (Y/N)

* Permutational ANOVA (PERMANOVA) followed by pairwise PERMANOVA with FDR correction (Benjamini-Hochberg method) if significant

```{r}
set.seed(6926)
```

## Archaea
### Fecal
Subset data
```{r}
meta.temp = metaA %>%
  filter(Type == "F") %>% 
  filter(AgeGroup %in% c("2w", "4w"))
```

```{r}
OTU.temp = metaA %>%
  select(Sample, Type, AgeGroup) %>% 
  filter(Type == "F") %>% 
  filter(AgeGroup %in% c("2w", "4w")) %>% 
  left_join(OTUA, by="Sample") %>% 
  select(-Sample, -Type, -AgeGroup)
```

**PERMANOVA** 
Bray-Curtis (structure)
```{r}
adonis(OTU.temp ~ ScourSample, data=meta.temp, strata=meta.temp$AgeGroup, method="bray", permutations=1000)
```

Jaccard (composition)
```{r}
adonis(OTU.temp ~ ScourSample, data=meta.temp, strata=meta.temp$AgeGroup, method="jaccard", permutations=1000)
```

## Bacteria
### Fecal
Subset data
```{r}
meta.temp = metaB %>%
  filter(Type == "F") %>% 
  filter(AgeGroup %in% c("2w", "4w"))
```

```{r}
OTU.temp = metaB %>%
  select(Sample, Type, AgeGroup) %>% 
  filter(Type == "F") %>% 
  filter(AgeGroup %in% c("2w", "4w")) %>% 
  left_join(OTUB, by="Sample") %>% 
  select(-Sample, -Type, -AgeGroup)
```

**PERMANOVA**  
Bray-Curtis (structure)
```{r}
adonis(OTU.temp ~ ScourSample, data=meta.temp, strata=meta.temp$AgeGroup, method="bray", permutations=1000)
```

Jaccard (composition)
```{r}
adonis(OTU.temp ~ ScourSample, data=meta.temp, strata=meta.temp$AgeGroup, method="jaccard", permutations=1000)
```

## Fungi
No fecal samples due to low abundance in young calves.
