---
title: "Individual OTU analyses"
subtitle: "GS01 Calf diets"
author: "Kim Dill-McFarland"
date: "version `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_float:
      collapsed: no
editor_options: 
  chunk_output_type: console
---
<script>
$(document).ready(function() {
  $items = $('div#TOC li');
  $items.each(function(idx) {
    num_ul = $(this).parentsUntil('#TOC').length;
    $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
  });
});
</script>

# Load data
Created in `1_data_manip.Rmd`
```{r}
set.seed(8325)
load("GS01.data.RData")
```

# Packages
```{r}
#Data manipulation
library(tidyverse)
```

# Functions
```{r }
source("functions/simper_pretty.R")
source("functions/R_krusk.R")
```

# Model 1: Diet and age (interaction only)
Diet:AgeGroup

* Similarity percentages (SIMPER)
    - up to 100% total cumulative SIMPER (`perc_cutoff=1`)
    - 1% minimum contribution from any single OTU (`low_cutoff = 'y', low_val=0.01`)
* Kruskal-Wallis with the Benjamini-Hochberg correction for multiple comparisons

Load in raw mothur-formatted taxonomy tables to work with `kruskal.pretty` function.
```{r message=FALSE}
A.taxonomy = read.table("data/taxonomy/GS01.Arc.final.fn.unique_list.0.03.cons.taxonomy", row.names=1, header=TRUE, sep="\t")

B.taxonomy = read.table("data/taxonomy/GS01.Bact.final.fn.unique_list.0.03.cons.taxonomy", row.names=1, header=TRUE, sep="\t")

F.taxonomy = read.table("data/taxonomy/GS01.Fun.final.fn.unique_list.0.03.cons.taxonomy", row.names=1, header=TRUE, sep="\t")
```

## Archaea
### Fecal
Subset data
```{r}
meta.temp = metaA %>%
  filter(Type == "F")
```

```{r}
OTU.temp = metaA %>%
  select(Sample, Type) %>% 
  filter(Type == "F") %>% 
  left_join(OTUA, by="Sample") %>% 
  select(-Sample, -Type)
```

**SIMPER**
```{r eval=FALSE}
simper.pretty(OTU.temp, meta.temp, c('DietAge'), perc_cutoff=1, low_cutoff = 'y', low_val=0.01, 'results/SIMPER/OTUAF')
```

**Kruskal-Wallis**
```{r eval=FALSE}
kruskal.pretty(OTU.temp, meta.temp, data.frame(read.csv('results/SIMPER/OTUAF_clean_simper.csv')), c('DietAge'), 'results/KruskalWallis/OTUAF', A.taxonomy)
```

### Rumen liquids
Subset data
```{r}
meta.temp = metaA %>%
  filter(Type == "L")
```

```{r}
OTU.temp = metaA %>%
  select(Sample, Type) %>% 
  filter(Type == "L") %>% 
  left_join(OTUA, by="Sample") %>% 
  select(-Sample, -Type)
```

**SIMPER**
```{r eval=FALSE}
simper.pretty(OTU.temp, meta.temp, c('DietAge'), perc_cutoff=1, low_cutoff = 'y', low_val=0.01, 'results/SIMPER/OTUAL')
```

**Kruskal-Wallis**
```{r eval=FALSE}
kruskal.pretty(OTU.temp, meta.temp, data.frame(read.csv('results/SIMPER/OTUAL_clean_simper.csv')), c('DietAge'), 'results/KruskalWallis/OTUAL', A.taxonomy)
```

### Rumen solids
Subset data
```{r}
meta.temp = metaA %>%
  filter(Type == "S")
```

```{r}
OTU.temp = metaA %>%
  select(Sample, Type) %>% 
  filter(Type == "S") %>% 
  left_join(OTUA, by="Sample") %>% 
  select(-Sample, -Type)
```

**SIMPER**
```{r eval=FALSE}
simper.pretty(OTU.temp, meta.temp, c('DietAge'), perc_cutoff=1, low_cutoff = 'y', low_val=0.01, 'results/SIMPER/OTUAS')
```

**Kruskal-Wallis**
```{r eval=FALSE}
kruskal.pretty(OTU.temp, meta.temp, data.frame(read.csv('results/SIMPER/OTUAS_clean_simper.csv')), c('DietAge'), 'results/KruskalWallis/OTUAS', A.taxonomy)
```

## Bacteria
### Fecal
Subset data
```{r}
meta.temp = metaB %>%
  filter(Type == "F")
```

```{r}
OTU.temp = metaB %>%
  select(Sample, Type) %>% 
  filter(Type == "F") %>% 
  left_join(OTUB, by="Sample") %>% 
  select(-Sample, -Type)
```

**SIMPER**
```{r eval=FALSE}
simper.pretty(OTU.temp, meta.temp, c('DietAge'), perc_cutoff=1, low_cutoff = 'y', low_val=0.01, 'results/SIMPER/OTUBF')
```

**Kruskal-Wallis**
```{r eval=FALSE}
kruskal.pretty(OTU.temp, meta.temp, data.frame(read.csv('results/SIMPER/OTUBF_clean_simper.csv')), c('DietAge'), 'results/KruskalWallis/OTUBF', A.taxonomy)
```

### Rumen liquids
Subset data
```{r}
meta.temp = metaB %>%
  filter(Type == "L")
```

```{r}
OTU.temp = metaB %>%
  select(Sample, Type) %>% 
  filter(Type == "L") %>% 
  left_join(OTUB, by="Sample") %>% 
  select(-Sample, -Type)
```

**SIMPER**
```{r eval=FALSE}
simper.pretty(OTU.temp, meta.temp, c('DietAge'), perc_cutoff=1, low_cutoff = 'y', low_val=0.01, 'results/SIMPER/OTUBL')
```

**Kruskal-Wallis**
```{r eval=FALSE}
kruskal.pretty(OTU.temp, meta.temp, data.frame(read.csv('results/SIMPER/OTUBL_clean_simper.csv')), c('DietAge'), 'results/KruskalWallis/OTUBL', A.taxonomy)
```

### Rumen solids
Subset data
```{r}
meta.temp = metaB %>%
  filter(Type == "S")
```

```{r}
OTU.temp = metaB %>%
  select(Sample, Type) %>% 
  filter(Type == "S") %>% 
  left_join(OTUB, by="Sample") %>% 
  select(-Sample, -Type)
```

**SIMPER**
```{r eval=FALSE}
simper.pretty(OTU.temp, meta.temp, c('DietAge'), perc_cutoff=1, low_cutoff = 'y', low_val=0.01, 'results/SIMPER/OTUBS')
```

**Kruskal-Wallis**
```{r eval=FALSE}
kruskal.pretty(OTU.temp, meta.temp, data.frame(read.csv('results/SIMPER/OTUBS_clean_simper.csv')), c('DietAge'), 'results/KruskalWallis/OTUBS', A.taxonomy)
```

## Fungi
### Fecal
Subset data
```{r}
meta.temp = metaF %>%
  filter(Type == "F")
```

```{r}
OTU.temp = metaF %>%
  select(Sample, Type) %>% 
  filter(Type == "F") %>% 
  left_join(OTUF, by="Sample") %>% 
  select(-Sample, -Type)
```

**SIMPER**
```{r eval=FALSE}
simper.pretty(OTU.temp, meta.temp, c('DietAge'), perc_cutoff=1, low_cutoff = 'y', low_val=0.01, 'results/SIMPER/OTUFF')
```

**Kruskal-Wallis**
```{r eval=FALSE}
kruskal.pretty(OTU.temp, meta.temp, data.frame(read.csv('results/SIMPER/OTUFF_clean_simper.csv')), c('DietAge'), 'results/KruskalWallis/OTUFF', A.taxonomy)
```

### Rumen liquids
Subset data
```{r}
meta.temp = metaF %>%
  filter(Type == "L")
```

```{r}
OTU.temp = metaF %>%
  select(Sample, Type) %>% 
  filter(Type == "L") %>% 
  left_join(OTUF, by="Sample") %>% 
  select(-Sample, -Type)
```

**SIMPER**
```{r eval=FALSE}
simper.pretty(OTU.temp, meta.temp, c('DietAge'), perc_cutoff=1, low_cutoff = 'y', low_val=0.01, 'results/SIMPER/OTUFL')
```

**Kruskal-Wallis**
```{r eval=FALSE}
kruskal.pretty(OTU.temp, meta.temp, data.frame(read.csv('results/SIMPER/OTUFL_clean_simper.csv')), c('DietAge'), 'results/KruskalWallis/OTUFL', A.taxonomy)
```

### Rumen solids
Subset data
```{r}
meta.temp = metaF %>%
  filter(Type == "S")
```

```{r}
OTU.temp = metaF %>%
  select(Sample, Type) %>% 
  filter(Type == "S") %>% 
  left_join(OTUF, by="Sample") %>% 
  select(-Sample, -Type)
```

**SIMPER**
```{r eval=FALSE}
simper.pretty(OTU.temp, meta.temp, c('DietAge'), perc_cutoff=1, low_cutoff = 'y', low_val=0.01, 'results/SIMPER/OTUFS')
```

**Kruskal-Wallis**
```{r eval=FALSE}
kruskal.pretty(OTU.temp, meta.temp, data.frame(read.csv('results/SIMPER/OTUFS_clean_simper.csv')), c('DietAge'), 'results/KruskalWallis/OTUFS', A.taxonomy)
```

## Select significant results
Read in Kruskal-Wallis results tables
```{r}
{# Set working directory for results files
require(plyr)
setwd("results/KruskalWallis/")
# List all Kruskal results files
file_list = list.files()

  # Read in all files and merge into 1 data frame
for(file in file_list){
  # if the merged dataset doesn't exist, create it
  if (!exists("KW_results")){
    # Read in 1st dataset
    KW_results = read.table(file, header=TRUE, sep="\t")
  # Add file name to data column to identify origin
    KW_results$origin = file
  }
   
  # if the merged dataset does exist, append to it
  if (exists("KW_results")){
    # Read in additional dataset
    temp_dataset =  read.csv(file, header=TRUE)
    # Add file name to data column to identify origin
    temp_dataset$origin = file
    # Merge datasets
    KW_results<-rbind.fill(KW_results, temp_dataset)
    # Remove original additional dataset
    rm(temp_dataset)
  }
}
# Return working directory to main project directory
setwd("../..")
}
```

Select significant (FDR P < 0.05) comparisons of interest.

* Diet A across ages
    - A2w_A4w, A4w_A8w, A8w_A1yr, A1yr_A2yr
* Diet B across ages
    - B2w_B4w, B4w_B8w, B8w_B1yr, B1yr_B2yr
* Diet C across ages
    - C2w_C4w, C4w_C8w, C8w_C1yr, C1yr_C2yr

* Age 2 weeks between diets
    - A2w_B2w, B2w_C2w, A2w_C2w
* Age 4 weeks between diets
    - A4w_B4w, B4w_C4w, A4w_C4w
* Age 8 weeks between diets
    - A8w_B8w, B8w_C8w, A8w_C8w
* Age 1 year between diets
    - A1yr_B1yr, B1yr_A1yr, C1yr_B1yr, A1yr_C1yr, C1yr_A1yr
* Age 2 years between diets
    - A2yr_B2yr, B2yr_A2yr, B2yr_C2yr, C2yr_B2yr, A2yr_C2yr

```{r}
KW_select = KW_results %>% 
  filter(fdr_krusk_p.val <= 0.05) %>% 
  filter(Comparison %in% c(
  "A2w_A4w", "A4w_A8w", "A8w_A1yr", "A1yr_A2yr",
  "B2w_B4w", "B4w_B8w", "B8w_B1yr", "B1yr_B2yr",
  "C2w_C4w", "C4w_C8w", "C8w_C1yr", "C1yr_C2yr",
         
  "A2w_B2w", "B2w_C2w", "A2w_C2w",
  "A4w_B4w", "B4w_C4w", "A4w_C4w",
  "A8w_B8w", "B8w_C8w", "A8w_C8w",
  "A1yr_B1yr", "B1yr_A1yr", "C1yr_B1yr", "A1yr_C1yr", "C1yr_A1yr",
  "A2yr_B2yr", "B2yr_A2yr", "B2yr_C2yr", "C2yr_B2yr", "A2yr_C2yr")) %>% 
  select(-X)

KW_select
```

Print significant results table.
```{r}
write.table(KW_select, file="results/KW_significant.csv", sep="\t")
```

# Model 2: Milk production efficiency
2yr samples only, averaged 3 consecutive rumen samples to avoid animal effects  

* Correlate OTUs to MPE or overall efficiency with Kendall’s correlations
* Strong correlations considered as <-0.7 or >0.7
* Minimum relative abundance of 0.5% in at least one sample

## Archaea
### Fecal
Subset data
```{r}
meta.temp = metaAave %>%
  filter(Type == "F" & AgeGroup == "2yr")
```

```{r}
OTU.temp = metaAave %>%
  select(Sample, Type, AgeGroup) %>% 
  filter(Type == "F" & AgeGroup == "2yr") %>% 
  left_join(OTUAave, by="Sample") %>% 
  select(-Sample, -Type, -AgeGroup)
```

**MPE**
```{r eval=FALSE}
write.table(cor(OTU.temp, meta.temp$MPE, method="kendall"), sep="\t", file="results/Kendall_milk/A.F.MPE.txt")
```

**Efficiency**
```{r eval=FALSE}
write.table(cor(OTU.temp, meta.temp$Effic, method="kendall"), sep="\t", file="results/Kendall_milk/A.F.effic.txt")
```

### Rumen liquids
Subset data
```{r}
meta.temp = metaAave %>%
  filter(Type == "L" & AgeGroup == "2yr")
```

```{r}
OTU.temp = metaAave %>%
  select(Sample, Type, AgeGroup) %>% 
  filter(Type == "L" & AgeGroup == "2yr") %>% 
  left_join(OTUAave, by="Sample") %>% 
  select(-Sample, -Type, -AgeGroup)
```

**MPE**
```{r eval=FALSE}
write.table(cor(OTU.temp, meta.temp$MPE, method="kendall"), sep="\t", file="results/Kendall_milk/A.L.MPE.txt")
```

**Efficiency**
```{r eval=FALSE}
write.table(cor(OTU.temp, meta.temp$Effic, method="kendall"), sep="\t", file="results/Kendall_milk/A.L.effic.txt")
```

### Rumen solids
Subset data
```{r}
meta.temp = metaAave %>%
  filter(Type == "S" & AgeGroup == "2yr")
```

```{r}
OTU.temp = metaAave %>%
  select(Sample, Type, AgeGroup) %>% 
  filter(Type == "S" & AgeGroup == "2yr") %>% 
  left_join(OTUAave, by="Sample") %>% 
  select(-Sample, -Type, -AgeGroup)
```

**MPE**
```{r eval=FALSE}
write.table(cor(OTU.temp, meta.temp$MPE, method="kendall"), sep="\t", file="results/Kendall_milk/A.S.MPE.txt")
```

**Efficiency**
```{r eval=FALSE}
write.table(cor(OTU.temp, meta.temp$Effic, method="kendall"), sep="\t", file="results/Kendall_milk/A.S.effic.txt")
```

## Bacteria
### Fecal
Subset data
```{r}
meta.temp = metaBave %>%
  filter(Type == "F" & AgeGroup == "2yr")
```

```{r}
OTU.temp = metaBave %>%
  select(Sample, Type, AgeGroup) %>% 
  filter(Type == "F" & AgeGroup == "2yr") %>% 
  left_join(OTUBave, by="Sample") %>% 
  select(-Sample, -Type, -AgeGroup)
```

**MPE**
```{r eval=FALSE}
write.table(cor(OTU.temp, meta.temp$MPE, method="kendall"), sep="\t", file="results/Kendall_milk/B.F.MPE.txt")
```

**Efficiency**
```{r eval=FALSE}
write.table(cor(OTU.temp, meta.temp$Effic, method="kendall"), sep="\t", file="results/Kendall_milk/B.F.effic.txt")
```

### Rumen liquids
```{r}
meta.temp = metaBave %>%
  filter(Type == "L" & AgeGroup == "2yr")
```

```{r}
OTU.temp = metaBave %>%
  select(Sample, Type, AgeGroup) %>% 
  filter(Type == "L" & AgeGroup == "2yr") %>% 
  left_join(OTUBave, by="Sample") %>% 
  select(-Sample, -Type, -AgeGroup)
```

**MPE**
```{r eval=FALSE}
write.table(cor(OTU.temp, meta.temp$MPE, method="kendall"), sep="\t", file="results/Kendall_milk/B.L.MPE.txt")
```

**Efficiency**
```{r eval=FALSE}
write.table(cor(OTU.temp, meta.temp$Effic, method="kendall"), sep="\t", file="results/Kendall_milk/B.L.effic.txt")
```

### Rumen solids
Subset data
```{r}
meta.temp = metaBave %>%
  filter(Type == "S" & AgeGroup == "2yr")
```

```{r}
OTU.temp = metaBave %>%
  select(Sample, Type, AgeGroup) %>% 
  filter(Type == "S" & AgeGroup == "2yr") %>% 
  left_join(OTUBave, by="Sample") %>% 
  select(-Sample, -Type, -AgeGroup)
```

**MPE**
```{r eval=FALSE}
write.table(cor(OTU.temp, meta.temp$MPE, method="kendall"), sep="\t", file="results/Kendall_milk/B.S.MPE.txt")
```

**Efficiency**
```{r eval=FALSE}
write.table(cor(OTU.temp, meta.temp$Effic, method="kendall"), sep="\t", file="results/Kendall_milk/B.S.effic.txt")
```

## Fungi
### Fecal
Subset data
```{r}
meta.temp = metaFave %>%
  filter(Type == "F" & AgeGroup == "2yr")
```

```{r}
OTU.temp = metaFave %>%
  select(Sample, Type, AgeGroup) %>% 
  filter(Type == "F" & AgeGroup == "2yr") %>% 
  left_join(OTUFave, by="Sample") %>% 
  select(-Sample, -Type, -AgeGroup)
```

**MPE**
```{r eval=FALSE}
write.table(cor(OTU.temp, meta.temp$MPE, method="kendall"), sep="\t", file="results/Kendall_milk/F.F.MPE.txt")
```

**Efficiency**
```{r eval=FALSE}
write.table(cor(OTU.temp, meta.temp$Effic, method="kendall"), sep="\t", file="results/Kendall_milk/F.F.effic.txt")
```

### Rumen liquids
Subset data
```{r}
meta.temp = metaFave %>%
  filter(Type == "L" & AgeGroup == "2yr")
```

```{r}
OTU.temp = metaFave %>%
  select(Sample, Type, AgeGroup) %>% 
  filter(Type == "L" & AgeGroup == "2yr") %>% 
  left_join(OTUFave, by="Sample") %>% 
  select(-Sample, -Type, -AgeGroup)
```

**MPE**
```{r eval=FALSE}
write.table(cor(OTU.temp, meta.temp$MPE, method="kendall"), sep="\t", file="results/Kendall_milk/F.L.MPE.txt")
```

**Efficiency**
```{r eval=FALSE}
write.table(cor(OTU.temp, meta.temp$Effic, method="kendall"), sep="\t", file="results/Kendall_milk/F.L.effic.txt")
```

### Rumen solids
Subset data
```{r}
meta.temp = metaFave %>%
  filter(Type == "S" & AgeGroup == "2yr")
```

```{r}
OTU.temp = metaFave %>%
  select(Sample, Type, AgeGroup) %>% 
  filter(Type == "S" & AgeGroup == "2yr") %>% 
  left_join(OTUFave, by="Sample") %>% 
  select(-Sample, -Type, -AgeGroup)
```

**MPE**
```{r eval=FALSE}
write.table(cor(OTU.temp, meta.temp$MPE, method="kendall"), sep="\t", file="results/Kendall_milk/F.S.MPE.txt")
```

**Efficiency**
```{r eval=FALSE}
write.table(cor(OTU.temp, meta.temp$Effic, method="kendall"), sep="\t", file="results/Kendall_milk/F.S.effic.txt")
```

## Select strong correlations
Read in Kendall results tables
```{r}
{# Set working directory for results files
setwd("results/Kendall_milk/")
# List all Kruskal results files
file_list = list.files()

  # Read in all files and merge into 1 data frame
for(file in file_list){
  # if the merged dataset doesn't exist, create it
  if (!exists("Kendall_milk_results")){
    # Read in 1st dataset
    Kendall_milk_results = read.table(file, header=TRUE, row.names=NULL, sep="\t")
  # Add file name to data column to identify origin
    Kendall_milk_results$origin = file
  }
   
  # if the merged dataset does exist, append to it
  if (exists("Kendall_milk_results")){
    # Read in additional dataset
    temp_dataset =  read.table(file, header=TRUE, row.names=NULL, sep="\t")
    # Add file name to data column to identify origin
    temp_dataset$origin = file
    # Merge datasets
    Kendall_milk_results<-rbind(Kendall_milk_results, temp_dataset)
    # Remove original additional dataset
    rm(temp_dataset)
  }
}
# Return working directory to main project directory
setwd("../..")
}
```

Select strong correlations <-0.7 or >0.7.
```{r}
Kendall_milk_select = Kendall_milk_results %>% 
  dplyr::rename(Kendall = V1, OTU=row.names) %>% 
  dplyr::filter(Kendall <= -0.7 | Kendall >=0.7) %>% 
  separate(origin, c("amplicon", "Type", "variable", "filetype"), by=".")
```

## Determine maximum relative abundances of OTUs.
### Archaea
```{r}
OTUA_2yr_abund = metaA %>%
  select(Sample, Animal, AgeGroup, Type) %>% 
  # Merge with OTU tables
  left_join(OTUA, by="Sample") %>% 
  # Use only 2 year samples
  filter(AgeGroup == "2yr") %>% 
  # Convert to percent relative abundance
  gather(key = OTU, value = count, -Sample, -Animal, -AgeGroup, -Type) %>% 
  group_by(Sample) %>%
  mutate(percent = count / sum(count) * 100) %>%

  # Find maximum abundance for each OTU
  group_by(OTU, Type) %>% 
  dplyr::summarize(max_perc = max(percent)) %>% 

  # Add column of amplicon type for later merging with other OTU tables
  mutate(amplicon="A")
```

### Bacteria
```{r}
OTUB_2yr_abund = metaB %>%
  select(Sample, Animal, AgeGroup, Type) %>% 
  # Merge with OTU tables
  left_join(OTUB, by="Sample") %>% 
  # Use only 2 year samples
  filter(AgeGroup == "2yr") %>% 
  # Convert to percent relative abundance
  gather(key = OTU, value = count, -Sample, -Animal, -AgeGroup, -Type) %>% 
  group_by(Sample) %>%
  mutate(percent = count / sum(count) * 100) %>%

  # Find maximum abundance for each OTU
  group_by(OTU, Type) %>% 
  dplyr::summarize(max_perc = max(percent)) %>% 

  # Add column of amplicon type for later merging with other OTU tables
  mutate(amplicon="B")
```

### Fungi
```{r}
OTUF_2yr_abund = metaF %>%
  select(Sample, Animal, AgeGroup, Type) %>% 
  # Merge with OTU tables
  left_join(OTUF, by="Sample") %>% 
  # Use only 2 year samples
  filter(AgeGroup == "2yr") %>% 
  # Convert to percent relative abundance
  gather(key = OTU, value = count, -Sample, -Animal, -AgeGroup, -Type) %>% 
  group_by(Sample) %>%
  mutate(percent = count / sum(count) * 100) %>%

  # Find maximum abundance for each OTU
  group_by(OTU, Type) %>% 
  dplyr::summarize(max_perc = max(percent)) %>% 

  # Add column of amplicon type for later merging with other OTU tables
  mutate(amplicon="F")
```

### Combine all
```{r}
OTU_2yr_abund = rbind(OTUA_2yr_abund, OTUB_2yr_abund)
OTU_2yr_abund = rbind(OTU_2yr_abund, OTUF_2yr_abund)
```

## Add abundances to strong correlations
```{r}
Kendall_milk_select_abund = Kendall_milk_select %>% 
  left_join(OTU_2yr_abund, by=c("OTU", "amplicon", "Type"))
```

Print significant results table.
```{r}
write.table(Kendall_milk_select_abund, file="results/Kendall_milk_significant.txt", sep="\t", row.names=FALSE)
```
