---
title: "SCOA analyses"
subtitle: "GS01 Calf diets"
author: "Kim Dill-McFarland"
date: "version `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_float:
      collapsed: no
editor_options: 
  chunk_output_type: console
---
<script>
$(document).ready(function() {
  $items = $('div#TOC li');
  $items.each(function(idx) {
    num_ul = $(this).parentsUntil('#TOC').length;
    $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
  });
});
</script>

# Load data
Created in `1_data_manip.Rmd`
```{r}
set.seed(8325)
load("GS01.data.RData")
```

# Packages
```{r}
#Beta-diversity analyses and plots
library(vegan)
#Data manipulation
library(tidyverse)
```

# Functions
```{r}
source("functions/Pairwise_adonis.R")
```

# Beta-diversity
## Model 1: Diet and age
Rumen liquid samples only  
Diet * AgeGroup + Random

* Permutational ANOVA (PERMANOVA) followed by pairwise PERMANOVA with FDR correction (Benjamini-Hochberg method) if significant

Subset data
```{r}
meta.temp = metaB %>%
  filter(Type == "L") %>% 
  arrange(Sample)
```

```{r}
SCOA.temp = SCOA %>%
  arrange(Sample) %>% 
  select(Succinate:Caproate) 
```

**PERMANOVA**
Bray-Curtis (structure)
```{r}
adonis(SCOA.temp ~ Diet*AgeGroup, data=meta.temp, method="bray", permutations=1000)
```

```{r}
pairwise.adonis(SCOA.temp, meta.temp$Diet, "bray", "fdr")
```

```{r}
pairwise.adonis(SCOA.temp, meta.temp$AgeGroup, "bray", "fdr")
```

```{r}
pairwise.adonis(SCOA.temp, meta.temp$DietAge, "bray", "fdr")
```

Jaccard (composition)
```{r}
adonis(SCOA.temp ~ Diet*AgeGroup, data=meta.temp, method="jaccard", permutations=1000)
```

```{r}
pairwise.adonis(SCOA.temp, meta.temp$Diet, "jaccard", "fdr")
```

```{r}
pairwise.adonis(SCOA.temp, meta.temp$AgeGroup, "jaccard", "fdr")
```

```{r}
pairwise.adonis(SCOA.temp, meta.temp$DietAge, "jaccard", "fdr")
```

## Model 2: Milk production efficiency
2yr rumen liquid samples only, averaged 3 consecutive rumen samples to avoid animal effects  
Effic + MPE + Random

* Permutational ANOVA (PERMANOVA) followed by pairwise PERMANOVA with FDR correction (Benjamini-Hochberg method) if significant

Subset data
```{r}
meta.temp = metaB %>%
  filter(Type == "L" & AgeGroup == "2yr") %>% 
  arrange(Sample)
```

```{r}
SCOA.temp = SCOA %>%
  filter(AgeGroup == "2yr") %>% 
  arrange(Sample) %>% 
  select(Succinate:Caproate) 
```

**PERMANOVA**
Bray-Curtis (structure)
```{r}
adonis(SCOA.temp ~ Effic + MPE, data=meta.temp, method="bray", permutations=1000)
```

Jaccard (composition)
```{r}
adonis(SCOA.temp ~ Effic + MPE, data=meta.temp, method="jaccard", permutations=1000)
```

# Specific SCOAs (mM and mF)
## Model 1: Diet and age
Diet * AgeGroup + Random

* ANOVAs with a Benjamini-Hochberg correction for multiple comparisons and pairwise between groups with TukeyHSD if significant 
* Concentraions and molar fractions tested separately

### Total VFA
acetate + propionate + isobutyrate + butyrate, isovalerate + 2-methylbutyrate + valerate

**Concentration**
```{r}
summary(aov(VFA_total_mM ~ Diet*AgeGroup, data=VFA))
```

```{r}
TukeyHSD(aov(VFA_total_mM ~ Diet*AgeGroup, data=VFA))
```

### Total branched
isobutyrate + isovalerate + 2-methylbutyrate

**Concentration**
```{r}
summary(aov(branch_mM ~ Diet*AgeGroup, data=VFA))
```

```{r}
TukeyHSD(aov(branch_mM ~ Diet*AgeGroup, data=VFA))
```

**Molar fraction**
```{r}
summary(aov(branch_mF ~ Diet*AgeGroup, data=VFA))
```

```{r}
TukeyHSD(aov(branch_mF ~ Diet*AgeGroup, data=VFA))
```

### Acetate
**Concentration**
```{r}
summary(aov(Acetate ~ Diet*AgeGroup, data=SCOA))
```

```{r}
TukeyHSD(aov(Acetate ~ Diet*AgeGroup, data=SCOA))
```

**Molar fraction**
```{r}
summary(aov(Acetate_mF ~ Diet*AgeGroup, data=VFA))
```

```{r}
TukeyHSD(aov(Acetate_mF ~ Diet*AgeGroup, data=VFA))
```

### Butyrate
**Concentration**
```{r}
summary(aov(Butyrate ~ Diet*AgeGroup, data=SCOA))
```

```{r}
TukeyHSD(aov(Butyrate ~ Diet*AgeGroup, data=SCOA))
```

**Molar fraction**
```{r}
summary(aov(Butyrate_mF ~ Diet*AgeGroup, data=VFA))
```

```{r}
TukeyHSD(aov(Butyrate_mF ~ Diet*AgeGroup, data=VFA))
```

### Ethanol
**Concentration**
```{r}
summary(aov(ETOH ~ Diet*AgeGroup, data=SCOA))
```

```{r}
TukeyHSD(aov(ETOH ~ Diet*AgeGroup, data=SCOA))
```

### Formate
**Concentration**
```{r}
summary(aov(Formate ~ Diet*AgeGroup, data=SCOA))
```

```{r}
TukeyHSD(aov(Formate ~ Diet*AgeGroup, data=SCOA))
```

### Isobutyrate
**Concentration**
```{r}
summary(aov(Isobutyrate ~ Diet*AgeGroup, data=SCOA))
```

```{r}
TukeyHSD(aov(Isobutyrate ~ Diet*AgeGroup, data=SCOA))
```

### Isovalerate + 2-methylbutyrate
**Concentration**
```{r}
summary(aov(iVal.2MB ~ Diet*AgeGroup, data=SCOA))
```

```{r}
TukeyHSD(aov(iVal.2MB ~ Diet*AgeGroup, data=SCOA))
```

### Lactate
**Concentration**
```{r}
summary(aov(Lactate ~ Diet*AgeGroup, data=SCOA))
```

```{r}
TukeyHSD(aov(Lactate ~ Diet*AgeGroup, data=SCOA))
```

### Propionate
**Concentration**
```{r}
summary(aov(Propionate ~ Diet*AgeGroup, data=SCOA))
```

```{r}
TukeyHSD(aov(Propionate ~ Diet*AgeGroup, data=SCOA))
```

**Molar fraction**
```{r}
summary(aov(Propionate_mF ~ Diet*AgeGroup, data=VFA))
```

```{r}
TukeyHSD(aov(Propionate_mF ~ Diet*AgeGroup, data=VFA))
```

### Succinate
**Concentration**
```{r}
summary(aov(Succinate ~ Diet*AgeGroup, data=SCOA))
```

```{r}
TukeyHSD(aov(Succinate ~ Diet*AgeGroup, data=SCOA))
```

### Valerate
**Concentration**
```{r}
summary(aov(Valerate ~ Diet*AgeGroup, data=SCOA))
```

```{r}
TukeyHSD(aov(Valerate ~ Diet*AgeGroup, data=SCOA))
```

**Molar fraction**
```{r}
summary(aov(Valerate_mF ~ Diet*AgeGroup, data=VFA))
```

```{r}
TukeyHSD(aov(Valerate_mF ~ Diet*AgeGroup, data=VFA))
```

### mM FDR correction
```{r}
#Diet
p.adjust(c(0.0125,0.00155,0.0788,0.41144,0.000158,0.00307,0.108,0.191,0.29548,0.63975,0.0217), method="fdr")
#AgeGroup
p.adjust(c(2E-16,2E-16,2E-16,0.02177,6.72E-10,2E-16,7.71E-11,2E-16,0.00379,0.47069,2E-16), method="fdr")
#Diet:AgeGroup
p.adjust(c(0.0000133,4.29E-08,0.000000898,0.00253,0.093772,0.25592,1.54E-09,3.21E-10,0.00123,0.00357,3.06E-11), method="fdr")
```

### mF FDR  correction
```{r}
#Diet
p.adjust(c(0.6317,0.537,0.58757,0.92762,0.1835), method="fdr")
#AgeGroup
p.adjust(c(4.43E-12,0.00000127,0.00203,1.24E-11,2E-16), method="fdr")
#Diet:AgeGroup
p.adjust(c(0.0506,0.101,0.49988,0.00031,0.0257), method="fdr")
```