---
title: "Co-variance of OTUs and SCOAs"
subtitle: "GS01 Calf diets"
author: "Kim Dill-McFarland"
date: "version `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_float:
      collapsed: no
editor_options: 
  chunk_output_type: console
---
<script>
$(document).ready(function() {
  $items = $('div#TOC li');
  $items.each(function(idx) {
    num_ul = $(this).parentsUntil('#TOC').length;
    $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
  });
});
</script>

# Load data
Created in `1_data_manip.Rmd`
```{r}
set.seed(8325)
load("GS01.data.RData")
```

# Packages
```{r Load packages}
#Beta-diversity analyses and plots
library(vegan)
#Data manipulation
library(tidyverse)
```

# Overall co-variation
Co-variation of the microbiota with SCOAs in rumen liquids was tested using Mantel tests for dissimilarity matrices.

Remove metadata from SCOA data to facilitate matrix calculations.
```{r}
SCOA.temp = SCOA %>%
  select(-Sample, -Diet, -AgeGroup, -TypeAll)
```

## Archaea
Subset data
```{r}
meta.temp = metaA %>%
  filter(Type == "L")
```

```{r}
OTU.temp = metaA %>%
  select(Sample, Type) %>% 
  filter(Type == "L") %>% 
  left_join(OTUA, by="Sample") %>% 
  select(-Sample, -Type)
```

**Mantel**
```{r}
mantel(
  vegdist(OTU.temp, method="bray"),
  vegdist(SCOA.temp, method="bray"),
  method="kendall", strata=meta.temp$AgeGroup, permutations=100)
```

## Bacteria
Subset data
```{r}
meta.temp = metaB %>%
  filter(Type == "L")
```

```{r}
OTU.temp = metaB %>%
  select(Sample, Type) %>% 
  filter(Type == "L") %>% 
  left_join(OTUB, by="Sample") %>% 
  select(-Sample, -Type)
```

**Mantel**
```{r}
mantel(
  vegdist(OTU.temp, method="bray"),
  vegdist(SCOA.temp, method="bray"),
  method="kendall", strata=meta.temp$AgeGroup, permutations=100)
```

## Fungi
Subset data
```{r}
meta.temp = metaF %>%
  filter(Type == "L")
```

```{r}
OTU.temp = metaF %>%
  select(Sample, Type) %>% 
  filter(Type == "L") %>% 
  left_join(OTUF, by="Sample") %>% 
  select(-Sample, -Type)
```

**Mantel**
```{r}
mantel(
  vegdist(OTU.temp, method="bray"),
  vegdist(SCOA.temp, method="bray"),
  method="kendall", strata=meta.temp$AgeGroup, permutations=100)
```

# Specific SCOAs and OTUs

* Correlate OTUs to SCOAs with Kendall’s correlations
* OTUs at least 0.05% relative abundance in one or more samples
* Averaged 3 consecutive rumen samples to avoid animal effects
* Rumen liquid samples only
* Strong correlations considered as <-0.7 or >0.7

Remove metadata from SCOA data to facilitate matrix calculations. Merge with mF VFA data
```{r}
SCOA.temp = SCOAave %>%
  full_join(VFAave, by="Sample") %>% 
  select(-Sample, -Diet.x, -Diet.y)
```

## Archaea
Subset data
```{r}
OTU.temp = metaAave %>%
  select(Sample, Type, AgeGroup) %>% 
  filter(Type == "L") %>% 
  left_join(OTUAave, by="Sample") %>% 
  select(-Sample, -Type, -AgeGroup)
```

**Kendall's correlation**
```{r eval=FALSE}
write.table(cor(OTU.temp, SCOA.temp, method="kendall"), sep="\t", file="results/Kendall_SCOA/A.L.OTUs.txt")
```

## Bacteria
Subset data
```{r}
OTU.temp = metaBave %>%
  select(Sample, Type, AgeGroup) %>% 
  filter(Type == "L") %>% 
  left_join(OTUBave, by="Sample") %>% 
  select(-Sample, -Type, -AgeGroup)
```

**Kendall's correlation**
```{r eval=FALSE}
write.table(cor(OTU.temp, SCOA.temp, method="kendall"), sep="\t", file="results/Kendall_SCOA/B.L.OTUs.txt")
```

## Fungi
Subset data
```{r}
OTU.temp = metaFave %>%
  select(Sample, Type, AgeGroup) %>% 
  filter(Type == "L") %>% 
  left_join(OTUFave, by="Sample") %>% 
  select(-Sample, -Type, -AgeGroup)
```

**Kendall's correlation**
```{r eval=FALSE}
write.table(cor(OTU.temp, SCOA.temp, method="kendall"), sep="\t", file="results/Kendall_SCOA/F.L.OTUs.txt")
```

# Specific calf SCOAs and OTUs

* Correlate OTUs to SCOAs present **only at 8 weeks** with Kendall’s correlations
    + Ethanol and lactate
* OTUs at least 0.05% relative abundance in one or more samples
* Averaged 3 consecutive rumen samples to avoid animal effects
* Rumen liquid samples only
* Strong correlations considered as <-0.7 or >0.7

Remove metadata from SCOA data to facilitate matrix calculations. Merge with mF VFA data
```{r}
SCOA.temp = SCOAave %>%
  full_join(VFAave, by="Sample") %>% 
  filter(grepl("8w", Sample)) %>% 
  select(-Sample, -Diet.x, -Diet.y) %>% 
  select(ETOH, Lactate)
```

## Archaea
Subset data
```{r}
OTU.temp = metaAave %>%
  select(Sample, Type, AgeGroup) %>% 
  filter(Type == "L" & AgeGroup == "8w") %>% 
  left_join(OTUAave, by="Sample") %>% 
  select(-Sample, -Type, -AgeGroup)
```

**Kendall's correlation**
```{r eval=FALSE}
write.table(cor(OTU.temp, SCOA.temp, method="kendall"), sep="\t", file="results/Kendall_SCOA/A.L.OTUs8w.txt")
```

## Bacteria
Subset data
```{r}
OTU.temp = metaBave %>%
  select(Sample, Type, AgeGroup) %>% 
  filter(Type == "L" & AgeGroup == "8w") %>% 
  left_join(OTUBave, by="Sample") %>% 
  select(-Sample, -Type, -AgeGroup)
```

**Kendall's correlation**
```{r eval=FALSE}
write.table(cor(OTU.temp, SCOA.temp, method="kendall"), sep="\t", file="results/Kendall_SCOA/B.L.OTUs8w.txt")
```

## Fungi
Subset data
```{r}
OTU.temp = metaFave %>%
  select(Sample, Type, AgeGroup) %>% 
  filter(Type == "L" & AgeGroup == "8w") %>% 
  left_join(OTUFave, by="Sample") %>% 
  select(-Sample, -Type, -AgeGroup)
```

**Kendall's correlation**
```{r eval=FALSE}
write.table(cor(OTU.temp, SCOA.temp, method="kendall"), sep="\t", file="results/Kendall_SCOA/F.L.OTUs8w.txt")
```

## Determine strong correlations
Read in Kendall results tables
```{r}
{# Set working directory for results files
  require(plyr)
  setwd("results/Kendall_SCOA/")
# List all Kruskal results files
file_list = list.files()

  # Read in all files and merge into 1 data frame
for(file in file_list){
  # if the merged dataset doesn't exist, create it
  if (!exists("Kendall_SCOA_results")){
    # Read in 1st dataset
    Kendall_SCOA_results = read.table(file, header=TRUE, row.names=NULL, sep="\t")
  # Add file name to data column to identify origin
    Kendall_SCOA_results$origin = file
  }
   
  # if the merged dataset does exist, append to it
  if (exists("Kendall_SCOA_results")){
    # Read in additional dataset
    temp_dataset =  read.table(file, header=TRUE, row.names=NULL, sep="\t")
    # Add file name to data column to identify origin
    temp_dataset$origin = file
    # Merge datasets
    Kendall_SCOA_results<-rbind.fill(Kendall_SCOA_results, temp_dataset)
    # Remove original additional dataset
    rm(temp_dataset)
  }
}
# Return working directory to main project directory
setwd("../..")
}
```

Select strong correlations <-0.7 or >0.7.
```{r}
Kendall_SCOA_select = Kendall_SCOA_results %>% 
  dplyr::rename(OTU=row.names) %>% 
  gather(key="SCOA", value="Kendall", Succinate:branch_mF) %>% 
  dplyr::filter(Kendall <= -0.7 | Kendall >=0.7) %>% 
  separate(origin, c("amplicon", "Type", "variable", "filetype"), by=".")
```

## Determine maximum relative abundances of OTUs.  
### Archaea
```{r}
OTUA_abund = metaA %>%
  select(Sample, Animal, AgeGroup, Type) %>% 
  # Merge with OTU tables
  left_join(OTUA, by="Sample") %>% 
  # Use only 2 year samples
  filter(AgeGroup %in% c("8w","1yr","2yr")) %>% 
  # Convert to percent relative abundance
  gather(key = OTU, value = count, -Sample, -Animal, -AgeGroup, -Type) %>% 
  group_by(Sample) %>%
  mutate(percent = count / sum(count) * 100) %>%

  # Find maximum abundance for each OTU
  group_by(OTU, Type) %>% 
  dplyr::summarize(max_perc = max(percent)) %>% 

  # Add column of amplicon type for later merging with other OTU tables
  mutate(amplicon="A")
```

### Bacteria
```{r}
OTUB_abund = metaB %>%
  select(Sample, Animal, AgeGroup, Type) %>% 
  # Merge with OTU tables
  left_join(OTUB, by="Sample") %>% 
  # Use only 2 year samples
  filter(AgeGroup %in% c("8w","1yr","2yr"))  %>% 
  # Convert to percent relative abundance
  gather(key = OTU, value = count, -Sample, -Animal, -AgeGroup, -Type) %>% 
  group_by(Sample) %>%
  mutate(percent = count / sum(count) * 100) %>%

  # Find maximum abundance for each OTU
  group_by(OTU, Type) %>% 
  dplyr::summarize(max_perc = max(percent)) %>% 

  # Add column of amplicon type for later merging with other OTU tables
  mutate(amplicon="B")
```

### Fungi 
```{r}
OTUF_abund = metaF %>%
  select(Sample, Animal, AgeGroup, Type) %>% 
  # Merge with OTU tables
  left_join(OTUF, by="Sample") %>% 
  # Use only 2 year samples
  filter(AgeGroup %in% c("8w","1yr","2yr"))  %>% 
  # Convert to percent relative abundance
  gather(key = OTU, value = count, -Sample, -Animal, -AgeGroup, -Type) %>% 
  group_by(Sample) %>%
  mutate(percent = count / sum(count) * 100) %>%

  # Find maximum abundance for each OTU
  group_by(OTU, Type) %>% 
  dplyr::summarize(max_perc = max(percent)) %>% 

  # Add column of amplicon type for later merging with other OTU tables
  mutate(amplicon="F")
```

### Combine all
```{r}
OTU_abund = rbind(OTUA_abund, OTUB_abund)
OTU_abund = rbind(OTU_abund, OTUF_abund)
```

## Add abundance to strong correlations
```{r}
Kendall_SCOA_select_abund = Kendall_SCOA_select %>% 
  left_join(OTU_abund, by=c("OTU", "amplicon", "Type")) %>% 
  distinct()
```

Print significant results table.
```{r}
write.table(Kendall_SCOA_select_abund, file="results/Kendall_SCOA_significant.txt", sep="\t", row.names=FALSE)
```
