---
title: "Figures 1-4"
subtitle: "GS01 Calf diets"
author: "Kim Dill-McFarland"
date: "version `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_float:
      collapsed: no
editor_options: 
  chunk_output_type: console
---
<script>
$(document).ready(function() {
  $items = $('div#TOC li');
  $items.each(function(idx) {
    num_ul = $(this).parentsUntil('#TOC').length;
    $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
  });
});
</script>

# Load data
Created in `1_data_manip.Rmd`
```{r}
set.seed(8325)
load("GS01.data.RData")
```

# Packages
```{r Load packages}
#Beta-diversity analyses and plots
library(vegan)
#Data manipulation
library(tidyverse)
# Plots
library(ggthemes)
library(RColorBrewer)
library(cowplot)
# HSD significance letters
library(agricolae)
```

# Figure 1
Total fecal and ruminal microbiota in dairy cows raised on different diets. 

## Define colors
```{r}
myPalette = colorRampPalette(brewer.pal(11, "Spectral"))
```

## Calculate nmds
**Archaea**
```{r results=FALSE}
AFnmds = OTUA %>% 
  dplyr::filter(grepl(".F", Sample)) %>% 
  select(-Sample) %>% 
  metaMDS(distance="bray", k=2, trymax=1000) %>% 
  scores(display="sites") %>% 
  as.data.frame()
AFnmds = OTUA %>% 
  dplyr::filter(grepl(".F", Sample)) %>% 
  select(Sample) %>% 
  cbind(AFnmds) %>% 
  mutate(amplicon="Arc")

ALnmds = OTUA %>% 
  dplyr::filter(grepl(".L", Sample)) %>% 
  select(-Sample) %>% 
  metaMDS(distance="bray", k=2, trymax=1000) %>% 
  scores(display="sites") %>% 
  as.data.frame()
ALnmds = OTUA %>% 
  dplyr::filter(grepl(".L", Sample)) %>% 
  select(Sample) %>% 
  cbind(ALnmds) %>% 
  mutate(amplicon="Arc")

ASnmds = OTUA %>% 
  dplyr::filter(grepl(".S", Sample)) %>% 
  select(-Sample) %>% 
  metaMDS(distance="bray", k=2, trymax=1000) %>% 
  scores(display="sites") %>% 
  as.data.frame()
ASnmds = OTUA %>% 
  dplyr::filter(grepl(".S", Sample)) %>% 
  select(Sample) %>% 
  cbind(ASnmds) %>% 
  mutate(amplicon="Arc")
```

**Bacteria**
```{r results=FALSE}
BFnmds = OTUB %>% 
  dplyr::filter(grepl(".F", Sample)) %>% 
  select(-Sample) %>% 
  metaMDS(distance="bray", k=2, trymax=1000) %>% 
  scores(display="sites") %>% 
  as.data.frame()
BFnmds = OTUB %>% 
  dplyr::filter(grepl(".F", Sample)) %>% 
  select(Sample) %>% 
  cbind(BFnmds) %>% 
  mutate(amplicon="Bac")

BLnmds = OTUB %>% 
  dplyr::filter(grepl(".L", Sample)) %>% 
  select(-Sample) %>% 
  metaMDS(distance="bray", k=2, trymax=1000) %>% 
  scores(display="sites") %>% 
  as.data.frame()
BLnmds = OTUB %>% 
  dplyr::filter(grepl(".L", Sample)) %>% 
  select(Sample) %>% 
  cbind(BLnmds) %>% 
  mutate(amplicon="Bac")

BSnmds = OTUB %>% 
  dplyr::filter(grepl(".S", Sample)) %>% 
  select(-Sample) %>% 
  metaMDS(distance="bray", k=2, trymax=1000) %>% 
  scores(display="sites") %>% 
  as.data.frame()
BSnmds = OTUB %>% 
  dplyr::filter(grepl(".S", Sample)) %>% 
  select(Sample) %>% 
  cbind(BSnmds) %>% 
  mutate(amplicon="Bac")
```

**Fungi**
```{r results=FALSE}
FFnmds = OTUF %>% 
  dplyr::filter(grepl(".F", Sample)) %>% 
  select(-Sample) %>% 
  metaMDS(distance="bray", k=2, trymax=1000) %>% 
  scores(display="sites") %>% 
  as.data.frame()
FFnmds = OTUF %>% 
  dplyr::filter(grepl(".F", Sample)) %>% 
  select(Sample) %>% 
  cbind(FFnmds) %>% 
  mutate(amplicon="Fun")

FLnmds = OTUF %>% 
  dplyr::filter(grepl(".L", Sample)) %>% 
  select(-Sample) %>% 
  metaMDS(distance="bray", k=2, trymax=1000) %>% 
  scores(display="sites") %>% 
  as.data.frame()
FLnmds = OTUF %>% 
  dplyr::filter(grepl(".L", Sample)) %>% 
  select(Sample) %>% 
  cbind(FLnmds) %>% 
  mutate(amplicon="Fun")

FSnmds = OTUF %>% 
  dplyr::filter(grepl(".S", Sample)) %>% 
  select(-Sample) %>% 
  metaMDS(distance="bray", k=2, trymax=1000) %>% 
  scores(display="sites") %>% 
  as.data.frame()
FSnmds = OTUF %>% 
  dplyr::filter(grepl(".S", Sample)) %>% 
  select(Sample) %>% 
  cbind(FSnmds) %>% 
  mutate(amplicon="Fun")
```

### Combine all
```{r}
all.nmds = rbind(AFnmds, ALnmds, ASnmds) %>% 
  rbind(BFnmds, BLnmds, BSnmds) %>% 
  rbind(FFnmds, FLnmds, FSnmds) %>% 
  
  gather(key=MDS, value=value, NMDS1:NMDS2) %>% 
  mutate(group=paste(amplicon, MDS, sep="_")) %>% 
  select(-amplicon, -MDS) %>% 
  
  group_by(Sample) %>% 
  spread(key=group, value=value) %>% 
  full_join(metaB[,c("Sample", "Animal", "AgeGroup", "Type", "Daysln", "Diet")], by="Sample") %>% 
  
  gather(key="key", value="value", Arc_NMDS1:Fun_NMDS2) %>% 
  separate(key, c("amplicon","MDS"), by="_") %>% 
  spread(key=MDS, value=value)
```

## Panels
**Main**
```{r warning=FALSE}
nmds_main = all.nmds %>% 

ggplot(aes(x=NMDS1, y=NMDS2)) +
  geom_point(aes(color=Daysln,
                 shape=Diet,
                 size=0.1),
             stroke=0.75) +
  scale_colour_gradientn(colours = myPalette(100), limits=c(2,7)) +
  scale_shape_manual(values = c(16,17,0)) +
  theme_base() +
  guides(color=FALSE, size=FALSE, shape=FALSE) +
  labs(x="", y="") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        plot.margin = margin(0.5,0.5,0,0, "cm")) +
  facet_wrap(amplicon~Type, scales = "free")
```

**Inset 1**
```{r warning=FALSE}
nmds_inset1 = all.nmds %>% 
  filter(amplicon=="Bac" & Type=="L" & AgeGroup %in% c("1yr","2yr")) %>% 

ggplot(aes(x=NMDS1, y=NMDS2)) +
  geom_point(aes(color=Daysln, shape=Diet, size=0.1), stroke=0.75) +
  scale_colour_gradientn(colours = myPalette(100), limits=c(2,7)) +
  scale_shape_manual(values = c(16,17,0)) +
  theme_base() +
  guides(color=FALSE, size=FALSE, shape=FALSE) +
  labs(x="", y="") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(), 
        panel.border = element_blank(), 
        plot.margin = margin(0.5,0.5,0,0, "cm"))
```

**Inset 2**
```{r warning=FALSE}
nmds_inset2 = all.nmds %>% 
  filter(amplicon=="Bac" & Type=="S" & AgeGroup %in% c("1yr","2yr")) %>% 

ggplot(aes(x=NMDS1, y=NMDS2)) +
  geom_point(aes(color=Daysln, shape=Diet, size=0.1), stroke=0.75) +
  scale_colour_gradientn(colours = myPalette(100), limits=c(2,7)) +
  scale_shape_manual(values = c(16,17,0)) +
  theme_base() +
  guides(color=FALSE, size=FALSE, shape=FALSE) +
  labs(x="", y="") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(), 
        plot.margin = margin(0.5,0.5,0,0, "cm"))
```

## Full plot
```{r warning=FALSE}
fig1 = ggdraw() +
  draw_plot(nmds_main, x = 0, y = 0, scale = 1) +
  draw_plot(nmds_inset1, x = 0.05, y = -0.1, scale = 0.18) +
  draw_plot(nmds_inset2, x = 0.4, y = -0.1, scale = 0.18)

fig1
```

```{r}
save_plot("figures/figure1.pdf", fig1,  base_height = 12, base_aspect_ratio = 0.9)
```

# Figure 2
Diversity of fecal and ruminal microbiota in cows raised on different diets.

## Define colors
```{r}
col_vec_age = c("#c9314c", "#f7894f", "#fddc88", "#72c7a4", "#4969ae")
```

## Merge data
```{r}
all.shan = metaA %>% 
  select(Sample, shannon) %>% 
  rename(Arc=shannon) %>% 
  
  full_join(metaF, by="Sample") %>% 
  select(Sample, Arc, shannon) %>% 
  rename(Fun=shannon) %>% 
  
  full_join(metaB, by="Sample") %>% 
  select(Sample, AgeGroup, Type, Arc, Fun, shannon) %>% 
  rename(Bac=shannon) %>% 
  
  gather(key="amplicon", value="shannon", Arc:Bac)
```

## Full plot
```{r warning=FALSE}
shan_plot = all.shan %>% 

ggplot() +
  geom_boxplot(aes(
    x=factor(AgeGroup, c("2w","4w","8w","1yr","2yr")), 
    y=shannon, 
    fill=factor(AgeGroup, c("2w","4w","8w","1yr","2yr")))) +
  
  scale_fill_manual(values = col_vec_age) +
  theme_base() +
  guides(fill=FALSE) +
  labs(x="", y="") +
  theme(plot.margin = margin(0.5,0.5,0,0, "cm")) +
  
  facet_wrap(amplicon~Type)

shan_plot
```

```{r}
save_plot("figures/figure2.pdf", shan_plot,  base_height = 9, base_aspect_ratio = 0.9)
```

## Determine significance letters
**Archaea**
```{r}
metaA %>% 
  mutate(group=paste(AgeGroup, Type, sep="_")) %>% 
  aov(shannon~group, .) %>% 
  HSD.test("group", group=TRUE) %>% 
  .[["groups"]]
```

**Bacteria**
```{r}
metaB %>% 
  mutate(group=paste(AgeGroup, Type, sep="_")) %>% 
  aov(shannon~group, .) %>% 
  HSD.test("group", group=TRUE) %>% 
  .[["groups"]]
```

**Fungi**
```{r}
metaF %>% 
  mutate(group=paste(AgeGroup, Type, sep="_")) %>% 
  aov(shannon~group, .) %>% 
  HSD.test("group", group=TRUE) %>% 
  .[["groups"]]
```

# Figure 3
Microbial taxa shared between calves and cows. 

1. In [mothur](https://www.mothur.org/wiki/Main_Page), merge AgeGroup+Diet groups
    + Example: `merge.groups(shared=Arc.shared, design=Arc.txt)`
2. In [mothur](https://www.mothur.org/wiki/Main_Page), get OTU list for each group
    + Example: `list.otulabels(shared=Bact.merge.shared, groups=group)`
3. Input OTU lists into  the Bioinformatics & Evolutionary Genomics [Venn diagram tool](http://bioinformatics.psb.ugent.be/webtools/Venn/)

# Figure 4
Calf diet impacts on weight gain and milk production.

## Define colors
```{r}
col_vec_diet = c("#7c590a", "#36a8e0", "#42871e")
```

## A. 2 days to weaning (8 weeks)
```{r}
calf_weight = weight %>%
  dplyr::filter(AgeExact<100) %>% 
  
  ggplot(aes(color=factor(Diet, c("A","C","B")))) + 
  stat_smooth(aes(x=AgeExact, y=Wt_norm), method="lm", se=FALSE) + 
  
  geom_point(aes(x=AgeExact, y=Wt_norm, shape=factor(Diet, c("A","C","B")))) +
  scale_shape(solid=FALSE, labels=c("Starter","Mix","Silage")) +
  scale_color_manual(values=col_vec_diet, labels=c("Starter","Mix","Silage")) +
  
  theme_base() +
  theme(legend.position="none") +
  labs(x="Age (days)", y="Weight gain (kg)")
```

## B. 8 weeks to first lactation
```{r}
cow_weight = weight %>%
  dplyr::filter(AgeExact>100) %>% 
  
  ggplot(aes(color=factor(Diet, c("A","C","B")))) + 
  stat_smooth(aes(x=AgeExact, y=Wt_norm), method="lm", se=FALSE) + 
  
  geom_point(aes(x=AgeExact, y=Wt_norm, shape=factor(Diet, c("A","C","B")))) +
  scale_shape(solid=FALSE, labels=c("Starter","Mix","Silage")) +
  scale_color_manual(values=col_vec_diet, labels=c("Starter","Mix","Silage")) +
  
  theme_base() +
  guides(color=guide_legend(title=""), shape=guide_legend(title="")) +
  labs(x="Age (days)", y="")
```

## C. Milk production
```{r}
milk_plot = metaBave %>%
  dplyr::filter(AgeGroup=="2yr" & Type=="F") %>% 
  select(Diet, MPE, Effic) %>% 
  gather(key=metric, value=value, MPE:Effic) %>% 
  
  ggplot() + 
  geom_boxplot(aes(x=factor(Diet, c("A","C","B")),
                   y=value,
                   fill=factor(Diet, c("A","C","B")))) +
  
  scale_fill_manual(values=col_vec_diet, labels=c("Starter","Mix","Silage")) +
  
  theme_base() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle=45, vjust = 0.01)) +
  scale_x_discrete(labels=c("A" = "Starter", "C" = "Mix", "B" = "Silage")) +
  labs(x="", y="Efficiency (MJ/kg DMI)") +
  facet_grid(~metric)
```

## Full plot
```{r warning=FALSE}
fig4 = plot_grid(calf_weight, cow_weight, milk_plot, nrow=1, rel_widths = c(3,3.9,3))

fig4
```

```{r}
save_plot("figures/figure4.pdf", fig4, base_width = 14, base_height = 4)
```
