---
title: "Supplemental figures 1-3"
subtitle: "GS01 Calf diets"
author: "Kim Dill-McFarland"
date: "version `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_float:
      collapsed: no
editor_options: 
  chunk_output_type: console
---
<script>
$(document).ready(function() {
  $items = $('div#TOC li');
  $items.each(function(idx) {
    num_ul = $(this).parentsUntil('#TOC').length;
    $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
  });
});
</script>

# Load data
Created in `1_data_manip.Rmd`
```{r}
set.seed(8325)
load("GS01.data.RData")
```

# Packages
```{r Load packages}
#Beta-diversity analyses and plots
library(vegan)
#Data manipulation
library(tidyverse)
# Plots
library(ggthemes)
#library(RColorBrewer)
library(cowplot)
# HSD significance letters
library(agricolae)
```

# Figure S1
Supplement intake (g/day) by diet from birth to weaning.

## Define colors
```{r}
col_vec_diet = c("#7c590a", "#36a8e0", "#42871e")
```

## A. Raw supplement intake
```{r}
intake_plot = ggplot(intake, aes(x = Day, y = Intake, 
                   color = factor(Diet, c("A","C","B")),
                   shape=factor(Diet, c("A","C","B")))) + 
  geom_point() + 
  scale_shape(solid=FALSE, labels=c("Starter","Mix","Silage")) +
  scale_color_manual(values=col_vec_diet, labels=c("Starter","Mix","Silage")) +
  
  theme_base() +
  labs(x="Age (days)", y="Feed intake (g)")  +
  guides(color=guide_legend(title=""), shape=guide_legend(title=""))
```

## B. Log supplement intake
```{r}
log_intake_plot = intake %>% 
  dplyr::filter(Intake >0) %>% 
  
ggplot(aes(x = Day, y = log(Intake), 
                   color = factor(Diet, c("A","C","B")),
                   shape=factor(Diet, c("A","C","B")))) + 
  geom_point() + 
  geom_smooth(method = "lm") +
  scale_shape(solid=FALSE, labels=c("Starter","Mix","Silage")) +
  scale_color_manual(values=col_vec_diet, labels=c("Starter","Mix","Silage")) +
  
  theme_base() +
  labs(x="Age (days)", y="Log feed intake")  +
  guides(color=guide_legend(title=""), shape=guide_legend(title=""))
```

## Full plot
```{r warning=FALSE}
figS1 = plot_grid(intake_plot, log_intake_plot, ncol=1)
figS1
```

```{r}
save_plot("figures/figureS1.pdf", figS1,  base_height = 9, base_aspect_ratio = 1.2)
```

# Figure S2
Diversity of fecal and ruminal microbiota in cows raised on different diets.

## Define colors
```{r}
col_vec_age = c("#c9314c", "#f7894f", "#fddc88", "#72c7a4", "#4969ae")
```

## Merge data
```{r}
all.chao = metaA %>% 
  select(Sample, chao) %>% 
  rename(Arc=chao) %>% 
  
  full_join(metaF, by="Sample") %>% 
  select(Sample, Arc, chao) %>% 
  rename(Fun=chao) %>% 
  
  full_join(metaB, by="Sample") %>% 
  select(Sample, AgeGroup, Type, Arc, Fun, chao) %>% 
  rename(Bac=chao) %>% 
  
  gather(key="amplicon", value="chao", Arc:Bac)
```

## Full plot
```{r warning=FALSE}
chao_plot = all.chao %>% 

ggplot() +
  geom_boxplot(aes(
    x=factor(AgeGroup, c("2w","4w","8w","1yr","2yr")), 
    y=chao, 
    fill=factor(AgeGroup, c("2w","4w","8w","1yr","2yr")))) +
  
  scale_fill_manual(values = col_vec_age) +
  theme_base() +
  guides(fill=FALSE) +
  labs(x="", y="") +
  theme(plot.margin = margin(0.5,0.5,0,0, "cm")) +
  
  facet_wrap(amplicon~Type, scales="free")

chao_plot
```

```{r warning=FALSE}
save_plot("figures/figureS2.pdf", chao_plot,  base_height = 9, base_aspect_ratio = 0.9)
```

## Determine significance letters
**Archaea**
```{r}
metaA %>% 
  mutate(group=paste(AgeGroup, Type, sep="_")) %>% 
  aov(chao~group, .) %>% 
  HSD.test("group", group=TRUE) %>% 
  .[["groups"]]
```

**Bacteria**
```{r}
metaB %>% 
  mutate(group=paste(AgeGroup, Type, sep="_")) %>% 
  aov(chao~group, .) %>% 
  HSD.test("group", group=TRUE) %>% 
  .[["groups"]]
```

**Fungi**
```{r}
metaF %>% 
  mutate(group=paste(AgeGroup, Type, sep="_")) %>% 
  aov(chao~group, .) %>% 
  HSD.test("group", group=TRUE) %>% 
  .[["groups"]]
```

# Figure S3
SCOA mM and mF by age and diet

## Define colors
```{r}
col_vec_age2 = c(rep("#fddc88",3), rep("#72c7a4",3), rep("#4969ae",3))
```

## Merge data
```{r warning=FALSE}
all.SCOA = SCOA %>% 
  select(-Animal, -Diet, -AgeGroup, -TypeAll) %>% 
  full_join(VFA, by = "Sample") %>% 
  
  rename(VFA_mM = VFA_total_mM) %>% 
  
  gather(key="key", value="value", c(Succinate:Caproate, VFA_mM:branch_mF)) %>% 
  
  mutate(group=key) %>% 
  separate(group, c("substrate", "units"), sep="_", fill="right") %>% 
  mutate(units = replace(units, is.na(units), "mM")) %>% 
  mutate(group = paste(AgeGroup, Diet, sep="")) %>% 
  mutate(group_ord = factor(group, c("8wA","8wC","8wB",
                                 "1yrA","1yrC","1yrB",
                                 "2yrA","2yrC","2yrB")))
```

## A. Concentration (mM)
```{r}
mM_plot = all.SCOA %>% 
  dplyr::filter(units == "mM") %>% 
  
  ggplot() +
  geom_boxplot(aes(x=group_ord, y=value, fill=group_ord)) +
  scale_fill_manual(values = col_vec_age2) +
  theme_base() +
  guides(fill=FALSE) +
  labs(x="", y="") +
  theme(plot.margin = margin(0.5,0.5,0,0, "cm")) +
  
  facet_wrap(units~substrate, ncol=3, scales="free_y")
```

## B. Molar fraction (mF)
```{r}
mF_plot = all.SCOA %>% 
  dplyr::filter(units == "mF") %>% 
  
  ggplot() +
  geom_boxplot(aes(x=group_ord, y=value, fill=group_ord)) +
  scale_fill_manual(values = col_vec_age2) +
  theme_base() +
  guides(fill=FALSE) +
  labs(x="", y="") +
  theme(plot.margin = margin(0.5,0.5,0,0, "cm")) +
  
  facet_wrap(units~substrate, ncol=3, scales="free_y")
```

## Full plot
```{r}
figS3 = plot_grid(mM_plot, mF_plot, ncol=1, rel_heights = c(2,1))
figS3
```

```{r}
save_plot("figures/figureS3.pdf", figS3,  base_height = 20, base_aspect_ratio = 0.5)
```

## Determine significance letters
### Function
```{r}
SCOA_letters = function(dat, vars, unit){
  require(tidyverse)
  require(agricolae)
  
  for(i in 1:length(vars)){
      test = dat %>% 
      dplyr::filter(substrate==vars[[i]] &
                    units==unit) %>% 
      aov(value~group, .) %>% 
      HSD.test("group", group=TRUE)
    
      letters = test$groups %>% 
        rownames_to_column() %>% 
        mutate(substrate=vars[[i]], units=unit)
    
  write.table(letters, file="figures/SCOA_letters.txt", append=TRUE, sep="\t", col.names = FALSE, row.names = FALSE)
    }
  
  y=read.table("figures/SCOA_letters.txt", header=FALSE, sep="\t", fill = TRUE)
  file.remove("figures/SCOA_letters.txt")
  colnames(y) = c("level", "value", "group", "substrate", "units")
  write.table(y, file="figures/SCOA_letters.txt", sep="\t", row.names = FALSE)
}
```

### Concentration  (mM)
```{r}
SCOA_letters(all.SCOA, vars=c("Acetate", "Butyrate", "ETOH", "Formate", "Isobutyrate", "iVal.2MB", "Lactate", "Propionate", "Succinate", "Valerate", "VFA"), unit="mM")
```

### Molar fraction  (mF)
```{r}
SCOA_letters(all.SCOA, vars=c("Acetate", "branch", "Butyrate", "Propionate", "Valerate"), unit="mF")
```

## Combine all
```{r}
letters = read.table("figures/SCOA_letters.txt", sep="\t", header=TRUE) %>% 
  filter(level != "level") %>% 
  select(-value) %>% 
  spread(key=level, value=group) %>% 
  write.table("figures/SCOA_letters_wide.txt", sep="\t")
```

